---
title: "Reading NC files from K disk, example (station 9: VR51 Korsen)"
output: 
  html_document:
    keep_md: true
---

Reading NettCDF (nc) files from the data repository  
`K:\Avdeling\214-Oseanografi\DATABASER\OKOKYST_2017\ncbase_OKOKYST`.   
  
Based on [this guide](http://geog.uoregon.edu/GeogR/topics/netCDF-read-ncdf4.html) for NTCD4.
  
Reads just two variables from one station as example (station 9: VR51 Korsen)

```{r}
library(ncdf4)

# Function that needs to be run for plotting using image() or image.plot()
transpose_for_image <- function(x) t(x[seq(nrow(x),1,-1),])
library(tidyverse)
```

```{r}
folder_data <- "K:/Avdeling/214-Oseanografi/DATABASER/OKOKYST_2017/ncbase_OKOKYST"
dir(folder_data)
```

Station list (from `print(ncin)` output, see below)
```
    22 global attributes:
        station01: VT3   Torbjoernskjaer   68936
        station02: VT10  Breiangen         31650
        station03: VT70  Bjoernafjorden    68910
        station04: VT69  Korsfjorden       68908
        station05: VT74  Maurangerfjorden  68913
        station06: VT53  Tveitneset        68911
        station07: VT16  Kyrkjeboe         68915
        station08: VT79  Naersnes          68914
        station09: VR51  Korsen            68871
        station10: VT71  Skinnbrokleia     63018
        station11: VT42  Korsfjorden       59644
        station12: VR52  Broemsneset       68869
        station13: VR31  Tilremsfjorden    68870
        station14: VT81  Alvenes           68950
        station15: VT82  Setsaa            68951
        station16: VT43  Kongsbakkneset    68937
        station17: VT28  Tjukkeneset       68933
        station18: VR54  Straumsfjorden    68938
        station19: VR7   Langfjordnes      68971
        station20: VR24  Tanafjorden       68972
        station21: VR21  Bugoeynes         68973
```

## Open file
```{r}
fn <- "VR51.nc"
ncin <- nc_open(paste0(folder_data, "/", fn))
# This makes a pretty long output:
# print(ncin)
# but it does 
```

## List dimensions and variables
Variables: `r names(ncin$var)`  
Dimensions: `r names(ncin$dim)`  

### Position of station
Longitude, latitude: `r ncvar_get(ncin,"lon")`, `r ncvar_get(ncin,"lat")`

### Dimensions
For depth, there are two types of dimensions:
  - for CTD variables (e.g. salinity), use 'depth'
  - for water sample varibles (e.g. Chl a, nutrients), use 'depth_nut'
#### Read dimensions
```{r}
time <- ncvar_get(ncin, "time")
depth <- ncvar_get(ncin, "depth")
depth_nut <- ncvar_get(ncin, "depth_nut")
```
#### Dimensions of time, depth, depth of nutrient samples
  - time: `r dim(time)`  
  - depth: `r dim(depth)`
  - depth_nut: `r dim(depth_nut)`  
  
#### Range of time, depth, depth of nutrient samples:
```{r}
range(time)
range(depth)
range(depth_nut)
```

#### Turn time into time object
```{r}
time <- as.POSIXct(time, tz = "UTC", origin = "1970-01-01")
```
Time range: `r range(time)`

## Salinity (as example of CTD variable) - read data
```{r}
salt <- ncvar_get(ncin, "salt")
dim(salt)
```
### Salinity, plot using fields
```{r}
fields::image.plot(transpose_for_image(salt))
```
### Make tidy data (for ggplot + anlysis)
```{r}
df <- data.frame(Depth = depth, salt)
colnames(df)[-1] <- paste0("t_", time)
df2 <- gather(df, "Time", "Salt", -Depth)
df2$Time <- lubridate::ymd_hms(sub("t_","",df2$Time))
head(df2, 3)
```

### Salinity, plot using ggplot
```{r}
ggplot(df2, aes(Time, Depth, fill = Salt)) + 
  geom_raster() +
  scale_fill_gradient2(low = "red", mid = "green", high = "blue", midpoint = 31) +
  scale_y_reverse()
```
### Salinity, plot profiles
```{r}
ggplot(df2, aes(Salt, Depth, color = factor(Time))) +
  geom_line() + 
  scale_y_reverse()
```
### Salinity, plot time series
```{r}
ggplot(df2 %>% filter(Depth %in% c(0,10,50)), aes(Time, Salt, color = factor(Depth))) +
  geom_line()
```

## Chlorophyll a (as example of a water sample variable)
```{r}
chla <- ncvar_get(ncin, "KlfA")
df <- data.frame(Depth = depth_nut, chla)        # using depth_nut instead of depth
colnames(df)[-1] <- paste0("t_", time)
df2 <- gather(df, "Time", "Chl a", -Depth)
df2$Time <- lubridate::ymd_hms(sub("t_","",df2$Time))
head(df2, 3)
```

### Chlorophyll a, image plot
```{r}
ggplot(df2, aes(Time, Depth, fill = `Chl a`)) + 
  geom_raster() +
  scale_fill_gradient2(low = "black", mid = "blue", high = "green", midpoint = 1) +
  scale_y_reverse()
```
### Chlorophyll a, time series plot
```{r}
ggplot(df2, aes(Time, `Chl a`, color = factor(Depth))) +
  geom_line()
```

