---
title: "Reading all nc files"
output: 
  html_document: 
    keep_md: yes
    toc: yes
---

## Libraries and functions
```{r}
library(ncdf4)
library(tidyverse)
library(ggmap)
library(readxl)

# Function that needs to be run for plotting using image() or image.plot()
transpose_for_image <- function(x) t(x[seq(nrow(x),1,-1),])

# Load function okokyst_read_nc()
source("02_Read_all_functions.R")

# Maps for used with ggplot
# map1 <- get_googlemap(c(lon=9, lat=62.5), zoom=4) %>% ggmap()
# map2 <- get_googlemap(c(lon=9, lat=62.5), zoom=5) %>% ggmap()

# Specify folder for nc files
folder_data <- "K:/Avdeling/214-Oseanografi/DATABASER/OKOKYST_2017/ncbase_OKOKYST"

```

## Get station attributes
```{r}
fn <- "K:/Avdeling/214-Oseanografi/DATABASER/OKOKYST_2017/OKOKYST_Hydrografi_Stasjoner_v5.xlsx"
df_stations <- read_excel(fn) %>% 
  rename(ProjectName = ProjcetName)
df_stations$StationName[9] <- "Korsfjorden_Hord"
df_stations$StationName[18] <- "Korsfjorden_Trønd"
```

## Test reads
### Test read 1
```{r}
# dir(folder_data)
df <- okokyst_read_nc("VT42_2011_2018.nc", "salt", ctd_variable = TRUE)

df %>% 
  arrange(Time, Depth) %>%
  ggplot(aes(salt, Depth, color = Time)) +
  geom_path() + 
  scale_y_reverse()

```

### Test read 2
```{r}
# debugonce(okokyst_read_nc)
df <- okokyst_read_nc("VT42_2011_2018.nc", "KlfA", ctd_variable = FALSE)

df %>% 
  arrange(Time, KlfA) %>%
  ggplot(aes(Time, KlfA, color = factor(Depth))) +
  geom_line()

```

## Read all variables from this station

### List variables  
```{r}
# Filenames
fn <- "VT42_2011_2018.nc"
ncin <- nc_open(paste0(folder_data, "/", fn))
names(ncin$var)

```

### Read dimensions
```{r}

time <- ncvar_get(ncin, "time")
depth <- ncvar_get(ncin, "depth")
depth_nut <- ncvar_get(ncin, "depth_nut")

#### Turn time into time object
time <- as.POSIXct(time, tz = "UTC", origin = "1970-01-01")

```


### Classify variables  
1. CTD variables (many depths, 455 in this case)  
2. Sample variables (from water samples, 10 depths in ths case)  
3. Secchi depth (ends up with the CTD variables in this case) 

```{r}
if (FALSE){
  x <- ncvar_get(ncin, "temp")
  dim(x)[1]  # 455
  x <- ncvar_get(ncin, "NO3")
  dim(x)[1]  # 11
}

get_depth_dim <- function(variable){
  x <- ncvar_get(ncin, variable)
  dim(x)[1]  # 1
}
# get_depth_dim("temp")

# 
vars <- names(ncin$var)
var_dims <- vars %>% map_int(get_depth_dim)

# MAke vars_ctd anf vars_sample  
vars_ctd <- vars[var_dims >= 20]
vars_sample <- vars[var_dims < 20]

cat("CTD variables\n")
vars_ctd

cat("Sample variables\n")
vars_sample

```


### Get CTD variables
```{r}
okokyst_read_nc_safe <- safely(okokyst_read_nc)
df_ctd_list <- vars_ctd %>% 
  map(~okokyst_read_nc_safe(
    "VT42_2011_2018.nc", 
    ., 
    ctd_variable = TRUE)
    )

names(df_ctd_list) <- vars_ctd
  
ok <- df_ctd_list %>% map_lgl(~is.null(.[[2]]))
ok_i <- which(ok)
# no 10 - secchi didn't work 

# check timed
if (FALSE){
  t1 <- df_ctd_list[["temp"]][[1]] %>% pull(Time) %>% unique()
  t2 <- df_ctd_list[["salt"]][[1]] %>% pull(Time) %>% unique()
  t3 <- df_ctd_list[["O2vol"]][[1]] %>% pull(Time) %>% unique()
  mean(t1 %in% t2)
  mean(t2 %in% t1)
  mean(t3 %in% t1)
}

# Start making df_ctd
df_ctd <- df_ctd_list[["temp"]][[1]]  
nrow(df_ctd)

# dput(vars_ctd)
vars_to_add <- c("salt", "turb", "fluor", "O2vol", "O2mg", "O2sat")

# Add variables to df_ctd, one at a time
for (var in vars_to_add){
  df_ctd <- df_ctd %>%
    full_join(df_ctd_list[[var]][[1]][c("Depth","Time", var)], 
              by = c("Depth","Time"))
  cat(var, ":", nrow(df_ctd), "\n")
}
# Add depths
for (var in c("depth1","depth2")){
  df_ctd <- df_ctd %>%
    left_join(df_ctd_list[[var]][[1]][c("Depth", var)], 
              by = c("Depth"))
  cat(var, ":", nrow(df_ctd), "\n")
}

# Rearrange columns
df_ctd <- df_ctd %>%
  select(Filename:Depth, depth1, depth2, everything())
df_ctd %>% names()
  
# Data density
names(df_ctd) %>% map_df(~data.frame(., mean(is.finite(df_ctd[[.]]))))

```

#### Check 1  
Max depth  
```{r}
df_ctd %>% 
  filter(is.finite(temp)) %>%
  group_by(Time) %>%
  summarise(Max_depth = max(depth1)) %>%
  ggplot(aes(Time, Max_depth)) +
  geom_line() +
  geom_point() +
  labs(title = "Max depth for temperature")

```

#### Check 2
```{r}
t1 <- df_ctd %>% pull(Time) %>% unique()
t1[1:5]

df_ctd %>%
  filter(Time %in% t1[1:20]) %>%
  arrange(Time, Depth) %>%
  ggplot(aes(temp, depth1, color = factor(Time))) +
  geom_path() + 
  scale_y_reverse()
  
df_ctd %>%
  filter(Time %in% t1[1:20]) %>%
  arrange(Time, Depth) %>%
  ggplot(aes(salt, depth1, color = factor(Time))) +
  geom_path() + 
  scale_y_reverse()
  
```

#### Profile 1  
No success    
```{r}
library(mgcv)

# Perform GAM and compute the predicted values from the GAM model
df_times <- df_ctd %>%
  count(Time) %>%
  mutate(Time_lag = lag(Time),
         Time_lead = lead(Time)) %>%
  mutate(
    Time_lag = 
      case_when(
        is.na(Time_lag) ~ Time - (Time_lead - Time),
        TRUE ~ Time_lag
      ),
    Time_lead = 
      case_when(
        is.na(Time_lead) ~ Time + (Time - Time_lag),
        TRUE ~ Time_lead
      ),
    Time_mid = Time_lag + (Time_lead - Time_lag)/2,
    Time_diff = Time_lead - Time_lag
  )
df_times

df <- df_ctd %>%
  filter(is.finite(temp)) %>%
  mutate(Depth_mid = (depth1 + depth2)/2) %>%
  left_join(df_times %>% select(-n))

# df %>% select(Time, Time_lag, Time_lead)

ggplot(df, aes(x = as.numeric(Time_mid), y = depth_mid, fill = temp, 
               height = depth1 - depth2, width = as.numeric(Time_diff))) +
  geom_tile() +
  scale_fill_gradient2(low = "blue4", mid = "green", high = "red2", midpoint = 10) +
  scale_y_reverse()

ggplot(df, aes(x = as.numeric(Time_mid), y = depth_mid, fill = temp, 
               height = depth1 - depth2)) +
  geom_tile() +
  scale_fill_gradient2(low = "blue4", mid = "green", high = "red2", midpoint = 10) +
  scale_y_reverse()

```


#### Profile 2
```{r, fig.width=9, fig.height=7}
library(mgcv)

# Perform GAM and compute the predicted values from the GAM model
df <- df_ctd %>%
  filter(is.finite(temp)) %>%
  mutate(Depth_mid = (depth1 + depth2)/2,
         Time_num = as.numeric(Time))
model <- gam(temp ~ te(Time_num, Depth_mid, k = 20), data = df)

df_smooth <- with(
  df,
  expand.grid(
    Time_num = seq(min(Time_num), max(Time_num), length = 150),
    Depth_mid = seq(min(Depth_mid), max(Depth_mid), length = 150)
  ))
df_smooth$Temp_gam <- predict.gam(model, df_smooth)
df_smooth$Time <- as.POSIXct(df_smooth$Time, origin = "1970-01-01", tz = "GMT")

#
# Set df_smooth to NA where there is no data
# df_smooth is NA if there are no data at that depth < 15 days away
#

# Max depth for data, 1 line per time (all times with data)
data_maxdepth <- df %>% 
  filter(is.finite(temp)) %>%
  group_by(Time_num) %>%
  summarise(Max_depth = max(depth1))
nrow(data_maxdepth)

# Times in the smoothed data
times_smooth <- sort(unique(df_smooth$Time_num))

# For times_smooth number i, return
#  maximum depth for all data within 15 days  
# global vars: df, times_smooth and data_maxdepth
get_maxdepth <- function(i, max_timediff = 15){
  df %>% 
    filter(is.finite(temp)) %>%
    count(Time_num) %>%
    mutate(Timediff = abs(times_smooth[i] - Time_num)/(24*3600))%>%
    select(-n) %>%
    # just add column from data_maxdepth, doon't need join as times should be identical 
    mutate(Max_depth = data_maxdepth$Max_depth) %>% 
    filter(Timediff <= max_timediff) %>%
    summarise(Max_depth = max(Max_depth)) %>%
    pull(Max_depth)
}
# Test
# get_maxdepth(2)

# Make maximum depth for every time in smooth
smooth_maxdepth <- data.frame(
  Time_num = times_smooth,
  Max_depth = seq_along(times_smooth) %>% map_dbl(get_maxdepth)
)
smooth_maxdepth

# Add maximum depth to smoothed data, and filter data so we keep only 
#   data < maximum depth
df_smooth <- df_smooth %>%
  left_join(smooth_maxdepth) %>%
  filter(Depth_mid <= Max_depth)

ggplot(df_smooth, aes(x = Time, y = Depth_mid)) +
  geom_tile(aes(fill = Temp_gam)) +
  # scale_fill_gradient2(low = "blue4", mid = "green", high = "red2", midpoint = 10) +
  scale_fill_gradientn(colours = fields::tim.colors(16)) +
  # viridis::scale_fill_viridis(option = "A") +
  geom_contour(aes(z = Temp_gam), binwidth = 1, color = "white") +
  # geom_point(data = df, aes(Time, Depth_mid), size = 0.1, color = "grey80") +
  scale_y_reverse() +
  theme_dark()



```

#### Profile 3   
Using okokyst_plot  
```{r, fig.width=9, fig.height=7}

t1 <- Sys.time()

okokyst_plot(data = df_ctd, varname = "temp", ctd_variable = TRUE)

t2 <- Sys.time()
t2-t1  # 15 sec

```

#### Profile 4   
Using okokyst_plot, linear  
```{r, fig.width=9, fig.height=7}

t1 <- Sys.time()

okokyst_plot(data = df_ctd, varname = "temp", ctd_variable = TRUE, linear = TRUE)

t2 <- Sys.time()
t2-t1   # 14.9 sec

```

#### Profile 5   
Using okokyst_plot, gam  
```{r, fig.width=9, fig.height=7}

t1 <- Sys.time()

okokyst_plot(data = df_ctd, varname = "temp", ctd_variable = TRUE, gam = TRUE, 
             gam_k = 20, nx = 150, ny = 150)

t2 <- Sys.time()
t2-t1  # 22 sec

# df_plot <- okokyst_make_plotdata(df, "salt")
# head(df_plot, 2)

```

```{r}
debugonce(okokyst_make_plotdata)
df_plot <- okokyst_make_plotdata(df, "salt", gam = TRUE)
head(df_plot, 2)

```

### Get Secchi data
```{r}

# Doesn't work
# df <- okokyst_read_nc_safe("VT42_2011_2018.nc", "secci", ctd_variable = TRUE)

X <- ncvar_get(ncin, "secchi")
dim(X)

df_secchi <- data.frame(
  time,
  secchi = X
)
                                
```


### Get water sample variables
```{r}
okokyst_read_nc_safe <- safely(okokyst_read_nc)
df_sample_list <- vars_sample %>% map(~okokyst_read_nc_safe("VT42_2011_2018.nc", 
                                                      ., 
                                                      ctd_variable = FALSE)
                                )

names(df_sample_list) <- vars_sample
  
ok <- df_sample_list %>% map_lgl(~is.null(.[[2]]))
ok_i <- which(ok)
# all worked  

# Start making df_sample
df_sample <- df_sample_list[["TotP"]][[1]]  
nrow(df_sample)

# dput(vars_sample)
vars_to_add <- c("PO4", "TotN", 
  "NO3", "NH4", "SiO2", "TSM", "KlfA", "CFYT", "CFYT1", "CFYT2", 
  "CFYT3", "FYT", "FYT1", "FYT2", "FYT3", "Sd")

# Add variables to df_sample, one at a time
for (var in vars_to_add){
  df_sample <- df_sample %>%
    full_join(df_sample_list[[var]][[1]][c("Depth","Time", var)], 
              by = c("Depth","Time"))
  cat(var, ":", nrow(df_sample), "\n")
}
# Add the rest
for (var in c("ProjectId", "StationId")){
  df_sample <- df_sample %>%
    left_join(df_sample_list[[var]][[1]][c("Depth", var)], 
              by = c("Depth"))
  cat(var, ":", nrow(df_sample), "\n")
}

# Rearraange
df_sample <- df_sample %>%
  select(Filename, ProjectId, StationId, Long:Sd)
df_sample %>% names()
  
# Data density
names(df_sample) %>% map_df(~data.frame(., mean(is.finite(df_sample[[.]]))))

```

#### Check 1  
Max depth  
```{r}
df_sample %>% 
  filter(is.finite(NO3)) %>%
  group_by(Time) %>%
  summarise(Max_depth = max(Depth)) %>%
  ggplot(aes(Time, Max_depth)) +
  geom_line() +
  geom_point() +
  labs(title = "Max depth for nitrate")

```


#### Check 2  
```{r}
t1 <- df_sample %>% pull(Time) %>% unique()

df_sample %>%
  filter(Time %in% t1[1:16 + 30]) %>%
  arrange(Time, Depth) %>%
  mutate(Time = factor(Time)) %>%
  ggplot(aes(NH4, Depth)) +
  geom_path() + 
  geom_point() + 
  scale_y_reverse() +
  facet_wrap("Time")

# Mimnimum values per date (LOQ in some dates, probably)
if (FALSE){
  df_sample %>%
    group_by(Time) %>%
    summarise_all(min, na.rm = TRUE) %>%
    View()
}

```

## QC
```{r}

openxlsx::write.xlsx(
  list(
    CTD_data = df_ctd,
    Secchi_data = df_secchi,
    Sample_data = df_sample
  ),
  "Data/02_Data_VT42.xlsx"
)

```


## Save   

```{r}

overwrite <- FALSE
overwrite <- TRUE    # backup will be made, see below

save_folder <- "Data"
save_filename <- "02_Data_VT42.xlsx"
save_fullpath <- paste0(save_folder, "/", save_filename)

file_exists <- save_filename %in% dir(save_folder)

if (file_exists & overwrite){
  cat("File exists and 'overwrite' = TRUE. File will be overwritten.\n")
  tempfile <- tempfile(pattern = "02_Data_", fileext = ".xlsx")
  cat("A temporary copy of the file will be saved with this name:\n")
  cat(tempfile, "\n")
  backup_saved <- file.copy(save_fullpath, tempfile)
} else if (file_exists & !overwrite) {
  cat("File exists and will not be overwritten. Set 'overwrite' to TRUE to change.\n")
  backup_saved <- FALSE
} else {
  backup_saved <- FALSE
}

  
if (overwrite & !backup_saved){
  cat("----------------------------------------------\nWas not able to save backup. File not overwritten.")
} else if (!file_exists | (overwrite & backup_saved)){
  openxlsx::write.xlsx(
    list(
      CTD_data = df_ctd,
      Secchi_data = df_secchi,
      Sample_data = df_sample
    ),
    save_fullpath
  )
  cat("----------------------------------------------\nFile saved.")
} 

# Delete backup files:
if (FALSE){
  fns <- dir(tempdir(), pattern = "02_Data_", full.names = TRUE)
  file.remove(fns)
}

```

