---
title: "Økokyst data - ChlA measured at 0 m vs 5 m"
output: 
  html_document: 
    keep_md: yes
    toc: yes
---

## Library
```{r}
library(tidyverse)
```

## Data
Saved in script 02
```{r}
df_chla <- readRDS("Data/02_df_chla.RData")
df_stations <- readRDS("Data/02_df_stations.RData")
```

## Make data set for plots
Pick 0 and 5 m and rearrange so 
```{r}
df_chla2 <- df_chla %>% 
  filter(Depth >= 0 & Depth <= 5) %>%
  spread(Depth, KlfA) %>%
  rename(Depth_00_m = `0`, Depth_05_m = `5`) %>%
  mutate(Abs_difference_00m_vs_05m = Depth_00_m - Depth_05_m,
         Perc_difference_00m_vs_05m = 100*(Depth_00_m - Depth_05_m)/Depth_05_m,
         StationName = factor(StationName, levels = df_stations$StationName))
df_chla2$StationCodeName  <- with(df_chla2, paste(StationCode, StationName))
# Checking (line 1 should be 2x line 2)
df_chla %>% filter(Depth <= 5) %>% nrow()
df_chla2 %>% nrow()
```

### Summary of data
```{r, echo=FALSE, warning=FALSE}
cat("Summary of data\n")
df_chla2 %>% 
  select(Depth_00_m, Depth_05_m, Abs_difference_00m_vs_05m, Perc_difference_00m_vs_05m) %>%
  summary()
cat("Abs_difference_00m_vs_05m, percentiles\n")
df_chla2$Abs_difference_00m_vs_05m %>% quantile(seq(0,1,0.2), na.rm = TRUE)
cat("Abs_difference_00m_vs_05m, percentiles of absolute values\n")
df_chla2$Abs_difference_00m_vs_05m %>% abs() %>% quantile(seq(0,1,0.2), na.rm = TRUE)
cat("Perc_difference_00m_vs_05m, percentiles\n")
df_chla2$Perc_difference_00m_vs_05m %>% quantile(seq(0,1,0.2), na.rm = TRUE)
cat("Perc_difference_00m_vs_05m, percentiles of absolute values\n")
df_chla2$Perc_difference_00m_vs_05m %>% abs() %>% quantile(seq(0,1,0.2), na.rm = TRUE)
```

## Absolute difference
### Absolute difference, 5 m vs 0 m {.tabset}

#### Full x and y scales
```{r, fig.height=7, fig.width=9, echo=FALSE, warning=FALSE}
absdiff_max <- df_chla2$Abs_difference_00m_vs_05m %>% abs() %>% max() %>% ceiling()
absdiff_breaks <- c(0,0.1,0.25,0.5,1,2, absdiff_max)
absdiff_colors <- c("green4","green","orange","orange3","red","red4")
gg <- ggplot(df_chla2, 
             aes(x = Depth_00_m, 
                 y = Depth_05_m, 
                 color = cut(abs(Abs_difference_00m_vs_05m), 
                             absdiff_breaks, include.lowest = TRUE)
                 )
             ) +
  geom_point() +
  scale_color_manual("Absolute\ndifference", values = absdiff_colors, na.value = "grey30") +
  geom_abline(aes(intercept = 0, slope = 1)) +
  facet_wrap(~StationCodeName)
gg
```

#### Restricted x and y scales (not showing all points)
```{r, fig.height=7, fig.width=9, echo=FALSE, warning=FALSE}
gg + 
  coord_cartesian(xlim = c(0, 3.5), ylim = c(0, 4))
```

### Absolute difference, time series {.tabset}
#### Full y scale
```{r, fig.width=9, echo=FALSE, warning=FALSE}
gg <- ggplot(df_chla2, aes(Time, Abs_difference_00m_vs_05m)) +
  geom_hline(aes(yintercept = 0)) +
  geom_line(color = "brown", size = 1) +
  facet_wrap(~StationCodeName)
gg
```

#### Restricted y scale
```{r, fig.width=9, echo=FALSE, warning=FALSE}
gg + coord_cartesian(ylim = c(-2.5, 2.5))
```

### Time series of Chl a at 5 m, absolute difference as colour {.tabset}
  * Co loured/connected dots: Chl a at 5 m  
  * Blue crosses: Chl a at 0 m
  
#### Full y scale
```{r, fig.height=7, fig.width=9, echo=FALSE, warning=FALSE}
gg <- ggplot(df_chla2, 
             aes(x = Time, 
                 y = Depth_05_m, 
                 color = cut(abs(Abs_difference_00m_vs_05m), 
                             absdiff_breaks, include.lowest = TRUE))
       ) +
  geom_line(color = "black") +
  geom_point() +
  geom_point(data = df_chla2, aes(Time, Depth_00_m), color = "blue", shape = 4) +
  scale_color_manual("Absolute\ndifference", values = absdiff_colors, na.value = "grey30") +
  facet_wrap(~StationCodeName)
gg
```

#### Restricted y scale
```{r, fig.width=9, echo=FALSE, warning=FALSE}
gg + coord_cartesian(ylim = c(0, 4.5))
```

## Percent difference
### Percent difference, 5 m vs 0 m {.tabset}

#### Full x and y scales
```{r, fig.height=7, fig.width=9, echo=FALSE, warning=FALSE}
percent_max <- df_chla2$Perc_difference_00m_vs_05m %>% abs() %>% max() %>% ceiling()
percent_breaks <- c(0,10,20,50,100,300, percent_max)
percent_colors <- c("green3","green","orange","orange3","red","red4")
gg <- ggplot(df_chla2, aes(Depth_00_m, Depth_05_m, 
                     color = cut(abs(Perc_difference_00m_vs_05m), percent_breaks, include.lowest = TRUE))
       ) +
  geom_point() +
  scale_color_manual("Percent\ndifference", values = percent_colors, na.value = "grey30") +
  geom_abline(aes(intercept = 0, slope = 1)) +
  facet_wrap(~StationCodeName)
gg
```


#### Restricted x and y scales (not showing all points)
```{r, fig.height=7, fig.width=9, echo=FALSE, warning=FALSE}
gg + 
  coord_cartesian(xlim = c(0, 3.5), ylim = c(0, 4))
```

### Percent difference, time series {.tabset}

#### Full y scale
```{r, fig.width=9, echo=FALSE, warning=FALSE}
ggplot(df_chla2, aes(Time, Perc_difference_00m_vs_05m)) +
  geom_line() +
  facet_wrap(~StationCodeName)

```

### Time series of Chl a at 5 m, percent difference as colour {.tabset}
  * Coloured/connected dots: Chl a at 5 m  
  * Blue crosses: Chl a at 0 m

#### Full y scale
```{r, fig.height=7, fig.width=9, echo=FALSE, warning=FALSE}
gg <- ggplot(df_chla2, aes(Time, Depth_05_m, 
                       color = cut(abs(Perc_difference_00m_vs_05m), percent_breaks, include.lowest = TRUE))
       ) +
  geom_line(color = "black") +
  geom_point() +
  geom_point(data = df_chla2, aes(Time, Depth_00_m), color = "blue", shape = 4) +
  scale_color_manual("Percent\ndifference", values = percent_colors, na.value = "grey30") +
  facet_wrap(~StationCodeName)
gg
```

#### Restricted y scale
```{r, fig.height=7, fig.width=9, echo=FALSE, warning=FALSE}
gg + 
  coord_cartesian(ylim = c(0, 4.5))
```

