---
title: "Read ?KOKYST data from NIVAbasen"
output: 
  html_document: 
    keep_md: yes
    toc: yes
---


- For ?kokyst 2017-2019 reports (Jan 2020)   
- If you want updated data, it can only be used for the stations/areas where NIVA collects data (Skagerrak)  
- For other areas, data stops in Feb. 2018 or something   
- However, for ?kokyst quality control, we used script 12 for Skagerrak and script 13 for Norskehavet S?r 1 and 2    
  
**Overview**
- Sections 1-3 makes data ready  
- Sections 4-5 are examples  
- Sections 6 and 7 (need only run 1-3 first) was used for checking whether TotN has gone down 20-30% after Eurofins took over TotN analyses    



## 1a. Libraries
```{r}
library(dplyr)
library(purrr)
library(ggplot2)
library(lubridate)  
library(readxl)

library(niRvana)
source("11_Read_from_Nivabasen_functions.R")

```

## 1b. Set username + password
```{r}

set_credentials()

```

## 2. Get projects

```{r}

df_projects <- get_projects()   # we call it 'df_projects' (the default name used by 'get_stations_from_project')

# Search for ?kokyst projects 
proj_names <- grep("?kokyst", df_projects$PROJECT_NAME, value = TRUE, ignore.case = TRUE)

# Filter for Skagerrak
proj_names <- grep("skagerrak",proj_names, value = TRUE, ignore.case = TRUE)

# Add extras (KYO + indre and ytre Oslofjord)
proj_names <- c(
  proj_names, 
  "Nordsj?en Nord",
  grep("indre oslo", df_projects$PROJECT_NAME, value = TRUE, ignore.case = TRUE), 
  grep("ytre oslo", df_projects$PROJECT_NAME, value = TRUE, ignore.case = TRUE), 
  grep("kyo", df_projects$PROJECT_NAME, value = TRUE, ignore.case = TRUE)
)


# get PROJECT_ID
proj_id <- df_projects %>% 
  filter(PROJECT_NAME %in% proj_names) %>%
  pull(PROJECT_ID)

# Get extra PROJECT_IDs from Steilene (to find Indre Oslofjord projects)
df <- get_nivabase_data("select * from NIVADATABASE.PROJECTS_STATIONS where STATION_NAME like 'Steilene%'") 
proj_id_extra <- df_projects %>% filter(PROJECT_ID %in% df$PROJECT_ID) %>% pull(PROJECT_ID)

proj_id <- c(proj_id, proj_id_extra) %>% unique()

#
```


## 3. Get stations
This gives you _many_ (1929!) stations, so we use the stations from the tender (next section) for the parts     
```{r}



df_stations <- get_nivabase_selection(
  "PROJECT_ID, STATION_ID, STATION_CODE, STATION_NAME",
  "PROJECTS_STATIONS",
  "PROJECT_ID",
  proj_id
)   

df_stations_geomid <- get_nivabase_selection(
  "STATION_ID, GEOM_REF_ID",
  "STATIONS",
  "STATION_ID",
  df_stations$STATION_ID
)   

df_stations_pos <- get_nivabase_selection(
  "SAMPLE_POINT_ID, LATITUDE, LONGITUDE",
  "SAMPLE_POINTS",
  "SAMPLE_POINT_ID",
  df_stations_geomid$GEOM_REF_ID,
  owner = "NIVA_GEOMETRY"
)   

df_stations <- df_stations %>% 
  left_join(df_stations_geomid) %>%
  left_join(df_stations_pos, by = c("GEOM_REF_ID" = "SAMPLE_POINT_ID"))

# df_stations %>% filter(STATION_CODE == "514")
df_stations <- df_stations %>%
  filter(LONGITUDE > 7.34 & LATITUDE < 60)

nrow(df_stations)  # 1247

# Make interactive map if you wish 
make_map <- FALSE
if (make_map){
  df <- df_stations %>% filter(STATION_NAME %in% "H??yfjorden")
  library(leaflet)
  leaflet() %>%
    addTiles() %>%  # Default OpenStreetMap map tiles
    addMarkers(lng = df$LONGITUDE, lat = df$LATITUDE,
               popup = paste(df$STATION_CODE, df$STATION_NAME))
}


```

### Hydrography stations in tender  
- Gives you Skagerrak sections  
- Can skip to section 6 after this      
```{r}

fn <- "K:/Prosjekter/Sj?vann/KYSTOVERV?KING ?KOKYST/KYSTOVERV?KING ?KOKYST 2017-2020/Stasjonsoversikt/?kokyststasjoner_tilbud2016_alle.xlsx"

st_hydr <- read_excel(fn, sheet = "hydrografi")

df_stations_hydrografi <- df_stations %>%
  filter(STATION_CODE %in% st_hydr$kode)

# Note! Only 7 stations found, 31 NOT found
st_hydr %>%
  filter(kode %in% df_stations_hydrografi$STATION_CODE) %>%
  nrow()
st_hydr %>%
  filter(!kode %in% df_stations_hydrografi$STATION_CODE) %>%
  nrow()

#  Check Bj?rnafjorden  
# 5,4742	60,1043
if (FALSE){
  df_stations %>% filter(
    LONGITUDE > 5.46 & LONGITUDE < 5.49 & 
      LATITUDE > 60.09 & LATITUDE < 60.12
  )
}

# Make interactive map if you wish 
make_map <- TRUE
make_map <- FALSE
if (make_map){
  df <- df_stations_hydrografi
  library(leaflet)
  leaflet() %>%
    addTiles() %>%  # Default OpenStreetMap map tiles
    addMarkers(lng = df$LONGITUDE, lat = df$LATITUDE,
               popup = paste(df$STATION_CODE, df$STATION_NAME))
}

```


## 4. Samples from one station    
Last date is 2.2.2018  
With "date visualisation" (could be better using label) 
```{r, fig.width=8, fig.height=2}

# Alt 1
df_station <- df_stations %>% 
  filter(STATION_NAME %in% "H??yfjorden")

# Alt 2
df_station <- df_stations %>% 
  filter(STATION_CODE %in% "VT67")

station_id <- unique(df_station$STATION_ID)

df_samples <- get_nivabase_selection(
  "WATER_SAMPLE_ID, STATION_ID, SAMPLE_DATE, DEPTH1, DEPTH2",
  "WATER_SAMPLES",
  "STATION_ID",
  station_id
)   

df <- tibble(Date = unique(df_samples$SAMPLE_DATE), Y = 0.5)


df %>%
  filter(year(Date) >= 2015) %>%
  ggplot() +
  geom_text(aes(Date, Y, label = Date), angle = 90, size = 2) +
  theme(axis.text = element_blank(),
        axis.title = element_blank())


```

## 5. Data from one date  
```{r}

unique(df_samples$SAMPLE_DATE) %>% tail(5)

sample_date <- df_samples %>%
  filter(floor_date(SAMPLE_DATE, "days") == ymd("2019-11-13")) %>%
  pull(SAMPLE_DATE) %>%
  unique()

df_data <- get_nivabase_selection(
  "WATER_SAMPLE_ID, METHOD_ID, VALUE, UNCERTAINTY, FLAG1, FLAG2, REMARK",
  "WATER_CHEMISTRY_VALUES",
  "WATER_SAMPLE_ID",
  subset(df_samples, SAMPLE_DATE %in% sample_date)$WATER_SAMPLE_ID 
)   

df_methods <- get_nivabase_selection(
  "METHOD_ID, NAME, UNIT, BASIS_ID",
  "METHOD_DEFINITIONS",
  "METHOD_ID",
  unique(df_data$METHOD_ID) 
)   

df_data <- df_data %>% 
  left_join(df_methods) %>%
  left_join(df_samples)

nrow(df_data)

  
```

### Plot data from this date  
```{r}
df_data %>%
  arrange(NAME, DEPTH1) %>%
  ggplot(aes(VALUE, DEPTH1)) + 
  geom_path() +
  geom_point(size = 0.5) +
  scale_y_reverse() +
  facet_wrap(vars(NAME), scales = "free_x")

```

### Get data for this station, 2016 - 2018  
Using function get_data_given_date()
```{r}

sample_dates <- df_samples %>%
  filter(SAMPLE_DATE >= ymd("2016-01-01")) %>%
  pull(SAMPLE_DATE) %>%
  unique()


dat <- get_data_given_date(sample_dates, df_samples, df_methods)

```

### Plot all profiles  
```{r}
dat %>%
  filter(NAME %in% "Nitritt + nitrat") %>%
  mutate(Date = as.character(round_date(SAMPLE_DATE, "day"))) %>%
  arrange(Date, DEPTH1) %>%
  ggplot(aes(VALUE, DEPTH1, color = Date)) + 
  geom_path() +
  geom_point(size = 0.5) +
  scale_y_reverse() 
  # facet_wrap(vars(Date))

```


## 6 Check Tot-N before/after 1. Nov. 2018  
- In order to compare before/after starting using Eurofins  
- Use all data from stations with data in 2019  
  
### Get all samples  
```{r}
# Test one station:
# df <- get_samples_onestation("VT67")
# df <- get_samples_onestation("VR51")
# df <- get_samples_onestation("VR51", depth1 = c(0,5,10,20,30))
# nrow(df)

station_codes <- unique(df_stations_hydrografi$STATION_CODE)

# Get list of data frames (one per station)  
df_samples_list <- station_codes %>% 
  map(
    get_samples_onestation,
    depth1 = c(0,5,10,20,30,50,100,200,300,400),     # nutrient depths + some  
    date_from = "01-01-2010")

# Check which have data  
not_ok <- df_samples_list %>% map_lgl(is.null) 

# Combine results
df_samples <- bind_rows(df_samples_list[!not_ok])


ggplot(df_samples, aes(x = SAMPLE_DATE, y = STATION_CODE)) +
  geom_point() +
  geom_vline(xintercept = ymd_hms("2014-01-01 00:00:00"), linetype = 2)


```

### Get all data  
```{r}  
sample_ids <- df_samples %>% 
  filter(SAMPLE_DATE > ymd("2000-01-01")) %>% 
  pull(WATER_SAMPLE_ID) %>% 
  unique()

df_data <- get_nivabase_selection(
  "WATER_SAMPLE_ID, METHOD_ID, VALUE, UNCERTAINTY, FLAG1, FLAG2, REMARK",
  "WATER_CHEMISTRY_VALUES",
  "WATER_SAMPLE_ID",
  sample_ids 
)   

df_methods <- get_nivabase_selection(
  "METHOD_ID, NAME, UNIT, BASIS_ID",
  "METHOD_DEFINITIONS",
  "METHOD_ID",
  unique(df_data$METHOD_ID) 
)   

df_data <- df_data %>% 
  left_join(df_methods, by = "METHOD_ID") %>%
  left_join(df_samples, by = "WATER_SAMPLE_ID")

nrow(df_data)  # 13202

table(df_data$STATION_CODE)


```

### Overwrite existing plot files?  
```{r}
overwrite_plots <- FALSE
```


### Plot TOTN

```{r, fig.width=8, fig.height=8}

gg <- df_data %>%
  filter(NAME %in% "Total nitrogen") %>%
  mutate(Depth = factor(DEPTH1)) %>%
  filter(STATION_CODE %in% c("VT2","VT3","VT10")) %>%
  ggplot(aes(SAMPLE_DATE, VALUE, color = Depth)) +
  geom_line() +
  geom_point() +
  facet_grid(vars(Depth), vars(STATION_CODE)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = -45, hjust = 0)) +
  geom_vline(xintercept = ymd_hms("2018-11-01 00:00:00"), linetype = 2) +
  labs(title = "Total N (fig. a)", y = "Total N")

if (overwrite_plots)
  ggsave("Figures/11_06_TotN_a.png", gg, width = 8, height = 8, dpi = 400)

gg

```

```{r, fig.width=8, fig.height=8}

gg <- df_data %>%
  filter(NAME %in% "Total nitrogen") %>%
  mutate(Depth = factor(DEPTH1)) %>%
  filter(STATION_CODE %in% c("VT65","VT66","VT67")) %>%  
  ggplot(aes(SAMPLE_DATE, VALUE, color = Depth)) +
  geom_line() +
  geom_point() +
  facet_grid(vars(Depth), vars(STATION_CODE)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = -45, hjust = 0)) +
  geom_vline(xintercept = ymd_hms("2018-11-01 00:00:00"), linetype = 2) +
  labs(title = "Total N (fig. b)", y = "Total N")

if (overwrite_plots)
  ggsave("Figures/11_06_TotN_b.png", gg, width = 8, height = 8, dpi = 400)

gg

```

### Calculate ratio inorg. N / TotN
```{r}
df_nitrogen <- df_data %>%
  select(STATION_CODE, SAMPLE_DATE, DEPTH1, NAME, VALUE) %>%
  filter(NAME %in% c("Total nitrogen", "Nitritt + nitrat", "Ammonium")) %>%
  tidyr::pivot_wider(names_from = NAME, values_from = VALUE) %>%
  mutate(Inorg_N = `Nitritt + nitrat` + `Ammonium`,
         Ratio_inorg_tot = Inorg_N/`Total nitrogen`) %>%
  arrange(STATION_CODE, SAMPLE_DATE, DEPTH1)

df_nitrogen  

```

### Plot ratio inorg. N / TotN

```{r, fig.width=8, fig.height=8}

gg <- df_nitrogen %>%
  filter(STATION_CODE %in% c("VT2","VT3","VT10")) %>%
  mutate(Depth = factor(DEPTH1)) %>%
  ggplot(aes(SAMPLE_DATE, Ratio_inorg_tot, color = Depth)) +
  geom_line() +
  geom_point() +
  facet_grid(vars(Depth), vars(STATION_CODE)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = -45, hjust = 0)) +
  geom_vline(xintercept = ymd_hms("2018-11-01 00:00:00"), linetype = 2) +
  labs(title = "Ratio inorganic N / total N (fig. a)", y = "Ratio inorganic N / total N")
  
if (overwrite_plots)
  ggsave("Figures/11_06_Nratio_a.png", gg, width = 8, height = 8, dpi = 400)

gg

```


```{r, fig.width=8, fig.height=8}

gg <- df_nitrogen %>%
  filter(STATION_CODE %in% c("VT65","VT66","VT67")) %>%
  mutate(Depth = factor(DEPTH1)) %>%
  ggplot(aes(SAMPLE_DATE, Ratio_inorg_tot, color = Depth)) +
  geom_line() +
  geom_point() +
  facet_grid(vars(Depth), vars(STATION_CODE)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = -45, hjust = 0)) +
  geom_vline(xintercept = ymd_hms("2018-11-01 00:00:00"), linetype = 2) +
  labs(title = "Ratio inorganic N / total N (fig. b)", y = "Ratio inorganic N / total N")
  
if (overwrite_plots)
  ggsave("Figures/11_06_Nratio_b.png", gg, width = 8, height = 8, dpi = 400)

gg

```

### Calculate ratio TOC / TotN  
Mainly data at VT3
```{r}
df_data %>%
  filter(NAME %in% c("L?st organisk karbon", "Partikul?rt organisk karbon")) %>%
  xtabs(~year(SAMPLE_DATE) + STATION_CODE, .)

df_toc_nitrogen <- df_data %>%
  filter(STATION_CODE %in% "VT3") %>%
  select(STATION_CODE, SAMPLE_DATE, DEPTH1, NAME, VALUE) %>%
  filter(NAME %in% c("Total nitrogen", "L?st organisk karbon", "Partikul?rt organisk karbon")) %>%
  tidyr::pivot_wider(names_from = NAME, values_from = VALUE) %>%
  mutate(TOC = `L?st organisk karbon` + `Partikul?rt organisk karbon`,
         Ratio_TOC_TotN= TOC/`Total nitrogen`) %>%
  arrange(STATION_CODE, SAMPLE_DATE, DEPTH1) 

df_toc_nitrogen  

```

### Plot   
```{r}
gg <- df_toc_nitrogen %>%
  mutate(Depth = factor(DEPTH1)) %>%
  filter(!is.na(Ratio_TOC_TotN)) %>%
  ggplot(aes(SAMPLE_DATE, Ratio_TOC_TotN, color = Depth)) +
  geom_line() +
  geom_point() +
  facet_grid(vars(Depth)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = -45, hjust = 0)) +
  geom_vline(xintercept = ymd_hms("2018-09-01 00:00:00"), linetype = 2) +
  labs(title = "Ratio TOC / total N", y = "Ratio TOC / total N")
  
if (overwrite_plots)
  ggsave("Figures/11_07_Nratio_b.png", gg, width = 8, height = 6, dpi = 400)

gg

```


## 7 Check longer time scales  

### Tables  
Using same df_data
```{r}
xtabs(~NAME + year(SAMPLE_DATE), df_data)

df_data %>% 
  filter(NAME %in% c("Total nitrogen", "Total nitrogen (Tot-N)", "TOTN")) %>%
  xtabs(~STATION_CODE + year(SAMPLE_DATE), .)

df_data %>% 
  filter(NAME %in% c("Total nitrogen", "Total nitrogen (Tot-N)", "TOTN")) %>%
  xtabs(~NAME + STATION_CODE, .)


```

```{r}
gg <- df_data %>%
  filter(NAME %in% "Total nitrogen") %>%
  mutate(Depth = factor(DEPTH1)) %>%
  filter(STATION_CODE %in% c("VT2","VT3","VT10")) %>%
  ggplot(aes(SAMPLE_DATE, VALUE, color = Depth)) +
  geom_line() +
  geom_point() +
  facet_grid(vars(Depth), vars(STATION_CODE)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = -45, hjust = 0)) +
  geom_vline(xintercept = ymd_hms("2018-11-01 00:00:00"), linetype = 2) +
  labs(title = "Total N (fig. a)", y = "Total N") +

```

```{r}
df_data %>%
  count(NAME, UNIT)

```

### TOT-N time series  
```{r}

df <- df_data %>%
  mutate(
    NAME2 = case_when(
      NAME %in% c("Total nitrogen", "Total nitrogen (Tot-N)", "TOTN") ~ "Total nitrogen",
      TRUE ~ NAME
    ),
    VALUE = case_when(
      UNIT %in% c("? mol/l", "?M") ~ VALUE/0.07139,
      TRUE ~ VALUE
    )
  ) %>%
  select(STATION_CODE, SAMPLE_DATE, DEPTH1, NAME, NAME2, VALUE) %>%
  filter(STATION_CODE %in% c("VT10","VT67")) %>%
  filter(NAME2 %in% "Total nitrogen") %>%
  mutate(Depth = factor(DEPTH1))

df_table1 <- df %>%
  filter(DEPTH1 >= 0 & DEPTH1 <= 10) %>%
  group_by(STATION_CODE, Year = year(SAMPLE_DATE)) %>%
  summarise(TotN = mean(VALUE) %>% round(0)) %>%
  arrange(STATION_CODE, Year)

df_table1


df_table2 <- df %>%
  filter(DEPTH1 >= 0 & DEPTH1 <= 10) %>%
  group_by(STATION_CODE, Year = year(SAMPLE_DATE), Month = month(SAMPLE_DATE)) %>%
  summarise(TotN = mean(VALUE) %>% round(0)) %>%
  pivot_wider(names_from = Month, values_from = TotN) %>%
  select(STATION_CODE, Year, as.character(1:12)) %>%
  arrange(STATION_CODE, Year)

df_table2

openxlsx::write.xlsx(list(Annual_average = df_table1,
                          Monthly_average = df_table2),
                     "Figures/11_07_TotN.xlsx")

```

### Plot  
```{r}

gg <- df %>%
  mutate(Month = month(SAMPLE_DATE)) %>%
  filter(STATION_CODE == "VT67" & DEPTH1 <= 30 & Month %in% 3:10) %>%
  ggplot(aes(SAMPLE_DATE, VALUE, color = Depth)) +
  geom_line() +
  geom_point() +
  facet_grid(vars(Depth), vars(Month)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = -45, hjust = 0)) +
  geom_vline(xintercept = ymd_hms("2018-11-01 00:00:00"), linetype = 2) +
  labs(title = "Total N (fig. a)", y = "Total N")


# ggsave("Figures/11_06_TotN_a.png", gg, width = 8, height = 8, dpi = 400)

 gg  

```


## 8 Other plots  
  
### Chl a  

```{r, fig.width=8, fig.height=8}
overwrite_plots <- TRUE 

#
# All stations (the 3 first), since 2016
# 
gg <- df_data %>%
  filter(NAME %in% c("Klorofyll a", "Klorofyll A")) %>%
  mutate(Depth = factor(DEPTH1)) %>%
  filter(STATION_CODE %in% c("VT2","VT3","VT10")) %>%
  filter(year(SAMPLE_DATE) >= 2016) %>%
  ggplot(aes(SAMPLE_DATE, VALUE, color = Depth)) +
  geom_line() +
  geom_point() +
  facet_grid(vars(Depth), vars(STATION_CODE)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = -45, hjust = 0)) +
  geom_vline(xintercept = ymd_hms("2018-11-01 00:00:00"), linetype = 2) +
  labs(title = "Chlorophyll a (fig. a)", y = "Chlorophyll a")

if (overwrite_plots)
  ggsave("Figures/11_08_Chla_a.png", gg, width = 8, height = 8, dpi = 400)

gg

#
# All stations (the 3 last), since 2016
# 
gg <- df_data %>%
  filter(NAME %in% c("Klorofyll a", "Klorofyll A")) %>%
  mutate(Depth = factor(DEPTH1)) %>%
  filter(STATION_CODE %in% c("VT65","VT66","VT67")) %>%  
  filter(year(SAMPLE_DATE) >= 2016) %>%
  ggplot(aes(SAMPLE_DATE, VALUE, color = Depth)) +
  geom_line() +
  geom_point() +
  facet_grid(vars(Depth), vars(STATION_CODE)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = -45, hjust = 0)) +
  geom_vline(xintercept = ymd_hms("2018-11-01 00:00:00"), linetype = 2) +
  labs(title = "Chlorophyll a (fig. b)", y = "Chlorophyll a")

if (overwrite_plots)
  ggsave("Figures/11_08_Chlb_b.png", gg, width = 8, height = 8, dpi = 400)

gg


#
# The two stations with longer time series, since 2010
# 
gg <- df_data %>%
  filter(NAME %in% c("Klorofyll a", "Klorofyll A")) %>%
  mutate(Depth = factor(DEPTH1)) %>%
  filter(STATION_CODE %in% c("VT10", "VT67")) %>%
  ggplot(aes(SAMPLE_DATE, VALUE, color = Depth)) +
  geom_line() +
  geom_point() +
  facet_grid(vars(Depth), vars(STATION_CODE)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = -45, hjust = 0)) +
  geom_vline(xintercept = ymd_hms("2018-11-01 00:00:00"), linetype = 2) +
  labs(title = "Chlorophyll a (fig. c)", y = "Chlorophyll a")

if (overwrite_plots)
  ggsave("Figures/11_08_Chla_c.png", gg, width = 8, height = 8, dpi = 400)


overwrite_plots <- FALSE 

gg

```




