---
title: "12_QA_2019_from_excel"
author: "DHJ"
date: "13 1 2020"
output:
  html_document:
    keep_md: true
    toc: true
    toc_float: true

---

# QA for CTDs - ØKOKYST 2019 report (January 2020)    
- Check Excel files from Bergen      
- Only CTDs
- See mail from Caroline Mengeot <caroline.mengeot@niva.no>, Sent: torsdag 9. januar 2020 14:47  

   
**Veileder:**  
- `Documents\artikler\veiledere etc\02_2013_klassifiserings-veileder_.pdf`  
- Chlorophyll: page 91.   
"I SørNorge (til Stadt) anbefales det at innsamlingen starter i februar og avsluttes ved utgangen av oktober."   
"Nord for Stadt anbefales det at innsamlingsperioden strekker seg fra mars til og med september."   
90% percentiles should be used   
  
- Also see calculation example on on page 187-189, and the comment on page 189:  
"Det gjøres oppmerksom på at dataprogrammene som beregner 90-persentil, gjerne benytter ulike metoder". We use the Excel method PERCENTILE.EXC, corresponding to type 6 quantiles in R  
  
- Nutrients: page 102   
Vinter er november til og med februar   
Sommer er juni til og med august  

  
```{r, echo=FALSE, results='hide'}

## 1. Libraries

library(dplyr)
library(purrr)
library(ggplot2)
library(lubridate)
library(readxl)
library(tidyr)
library(RColorBrewer)

# library(niRvana)

source("13_QA_CTD_2019_functions.R")

# RColorBrewer::display.brewer.all()
```


## 1. Folders and file information 

```{r, results='hide'}

### Folders

basefolder <- "K:/Avdeling/214-Oseanografi/DATABASER/OKOKYST_2017/"

# Norskehavet Sør I
folder1 <- "OKOKYST_NH_Sor1_RMS/xlsbase/TilAquamonitor"

# Norskehavet Sør II
folder2 <- "OKOKYST_NH_Sor2_Aquakompetanse/xlsbase/TilAquamonitor"
folder3 <- "OKOKYST_NH_Sor2_MagneAuren/xlsbase/TilAquamonitor"
folder4 <- "OKOKYST_NH_Sor2_SNO/xlsbase/TilAquamonitor"


```


### Check file names, sheet names and variable names   
Makes 'fileinfo' which contains all of this info (one line per file)  
```{r, echo=FALSE}
 
folder <- folder1
# 1. Find sheet names here: (can be commented out afterwards)
# sheets_in_folder(folder)                                 
# 2. Enter the sheet names here:  
f1 <- vars_in_folder(folder, c("data","data","Data"))       

folder <- folder2
# sheets_in_folder(folder)
f2 <- vars_in_folder(folder, c("Rådata","Rådata","Data"))

folder <- folder3
# sheets_in_folder(folder)
f3 <- vars_in_folder(folder, c("Data","Rådata","Data"))

folder <- folder4
# sheets_in_folder(folder)
f4 <- vars_in_folder(folder, c("Data","Rådata","Data","Data"))

fileinfo <- bind_rows(f1, f2, f3, f4)

fileinfo

```

```{r check_Variables, echo=FALSE, results='hide'}

### Check which variables are in each file  

files_lacking_var <- function(variablename, fileinformation = fileinfo){
  data.frame(
    Variable = variablename, 
    No_of_files_lacking = fileinformation %>%
      filter(!grepl(variablename, Variables)) %>% 
      nrow(),
    stringsAsFactors = FALSE
  )
}

# Variables of first file:
# i <- 1
# fn <- with(fileinfo[1,], paste0(basefolder, Folder, "/", File))
# read_excel(fn, sheet = fileinfo[1,"Sheet"], n_max = 0) %>% names() %>% dput()

#
# Number of 
c("StationId", "StationCode", "Date", "Depth1", 
  "Saltholdighet", "Temperatur", "Oksygen", "Oksygenmetning", "Metode") %>% map_df(files_lacking_var)

#
# 9 files contains StationCode, the rest have StationId
#
data.frame(
  StationCode = grepl("StationCode", fileinfo$Variables),
  StationId = grepl("StationId", fileinfo$Variables)
)


```

```{r, echo=FALSE, results='hide'}

### Get stations  
# Make 'fileinfo_stations'   
# - has one row per file and station  
# - deleting one station from one file (just 1 observation)

# Test
# i <- 1
# fn <- with(fileinfo[i,], paste0(basefolder, Folder, "/", File))
# dat <- read_excel_droplines(fn, sheet = fileinfo$Sheet[i])

#
# Makes a data frame for a given file, with one row per station
#
# NOTE: Global variable: 'fileinfo'
get_stations <- function(file_number){
  i <- file_number
  fn <- with(fileinfo[i,], paste0(basefolder, Folder, "/", File))
  dat <- read_excel_droplines(fn, sheet = fileinfo$Sheet[i])
  if ("StationCode" %in% names(dat)){
    df <- dat %>% count(StationCode)
      StationCode_var = "StationCode"
  } else {
    df <- dat %>% count(StationId)
    StationCode_var = "StationId"
    names(df)[1] <- "StationCode"
  }
  df$StationCode <- as.character(df$StationCode)
  data.frame(
    Folder = fileinfo$Folder[i],
    File = fileinfo$File[i],
    Sheet = fileinfo$Sheet[i],
    StationCode_var = StationCode_var,
    File_no = i,
    df,
    stringsAsFactors = FALSE
  )
}

# Test
# get_stations(1)
# get_stations(4)

fileinfo_stations <- seq_len(nrow(fileinfo)) %>% map_df(get_stations)

# Deletes staton with 1 observation
fileinfo_stations <- 
  fileinfo_stations %>% filter(n >= 5)

fileinfo_stations

```

## 2. Plot casts

```{r, echo=FALSE, results='hide'}

# Data from file 1  
i <- 1
fn <- with(fileinfo[i,], paste0(basefolder, Folder, "/", File))
dat <- read_excel_droplines(fn, sheet = fileinfo[i,"Sheet"])

#
### Time series plots, test
#
df <- dat %>%
  filter(StationCode == "VR51") %>%
  mutate(Depth = case_when(
    Depth1 == 0 ~ 0,
    Depth1 > 0 ~ round((Depth1+Depth2)/2, 0) 
  ))

# pie(rep(1,9), col = rainbow(n = 9, end = 0.75, v = 0.8))
depths <- c(0, 1, 5, 10, 20, 50, 100, 200, max(df$Depth)-10)
colors <- rainbow(n = 9, end = 0.75, v = 0.8)
colors <- c("#18A439", "#6D8F00", "#8B7800", "#9A6100", "#9E4B3F", "#983861", "#882F77", "#6B3480", "#3C417E")
names(colors) <- depths
colors

gg <- df %>%
  filter(Depth %in% depths) %>%
  mutate(Depth = factor(Depth)) %>%
  ggplot(aes(Date, Saltholdighet, group = Depth, color = Depth)) +
  geom_line() + geom_point(size = 1) +
  scale_color_manual(values = colors) +
  theme_bw()

# library(colorspace)

```


```{r, echo=FALSE, results='hide'}

#
# Plot profile and time series plot, example    
#

if (FALSE){
  plot_ctdprofile_station("VR51", dat, "Saltholdighet")
  plot_timeseries_station("VR51", dat, "Saltholdighet")
}

```

```{r, echo=FALSE, results='hide'}

### Test contour plot   
# See script 06  
# Didn't work that well


if (FALSE){
  source("02_Read_all_functions.R")
  dat %>%
    filter(StationCode == "VR51") %>%
    rename(Depth = Depth1, Time = Date) %>%
    okokyst_plot(., "Saltholdighet", ctd_variable = TRUE, binwidth = 20)
}

```



```{r functions, echo = FALSE}

### Functions

#
# Read data based on given row number in 'fileinfo_stations' file
#
# Note: 'fileinfo_stations' is global
#
read_data_fileinfoline <- function(i){
    fn <- with(fileinfo_stations[i,], 
               paste0(basefolder, Folder, "/", File))
    dat <- read_excel_droplines(fn, 
                                sheet = fileinfo_stations$Sheet[i])
    dat
}

#
# Plot all casts, function
# Note: 'fileinfo_stations' is global
#
plot_cast_all <- function(){
  for (i in seq_len(nrow(fileinfo_stations))){
    dat <- read_data_fileinfoline(i)
    file_no <- fileinfo_stations$File_no[i]
    plot_cast_station(fileinfo_stations$StationCode[i], 
                            dat, 
                            titletext = paste0("Plot ", i, ", File ", file_no)
    )
  }
}


#
# Plot all time series plots, function      
#
# Note: 'fileinfo_stations' is global
#
plot_timeseries_all <- function(variable){
  for (i in seq_len(nrow(fileinfo_stations))){
    dat <- read_data_fileinfoline(i)
    file_no <- fileinfo_stations$File_no[i]
    plot_timeseries_station(fileinfo_stations$StationCode[i], 
                            dat, 
                            variable, 
                            titletext = paste0("Plot ", i, ", File ", file_no, " -")
    )
  }
}

#
# Plot all profiles, function      
#
# Note: 'fileinfo_stations' is global
#
plot_ctdprofile_all <- function(variable){
  for (i in seq_len(nrow(fileinfo_stations))){
    dat <- read_data_fileinfoline(i)
    file_no <- fileinfo_stations$File_no[i]
    plot_ctdprofile_station(fileinfo_stations$StationCode[i], 
                            dat, 
                            variable, 
                            titletext = paste0("Plot ", i, ", File ", file_no, " -")
    )
  }
}

# plot_ctdprofile_all("Saltholdighet")

```

Plot all casts  
- Also check max depth    
**Notes**  
- 14. Dec 2018 (VR51) seems not to have reached the bottom (or was taken in the wrong place)  
- 14. Dec 2018 (VR51) also occurs in two files (files named "2018" and "2019"), which may lead be aware so the resukts are not used twice  
```{r}

plot_cast_all()

```

## 3. Plot parameters

### Plot salinity

```{r}

plot_timeseries_all("Saltholdighet")
plot_ctdprofile_all("Saltholdighet")

```

### Plot temperature  
Seems suspicious (see data in next chunk):  
- File 5, Vr52 26.4.2018 - temp + salinity down to 2.5 meters (3 measurements)   
- File 12, station 68870, 2018-08-21 - temp + salinity down to 0.5 meters  

```{r}

plot_timeseries_all("Temperatur")
plot_ctdprofile_all("Temperatur")

```

#### Checking some data  
- File 5, Vr52 26.4.2018
```{r}
# File 3, Vr52 26.4.2018
dat <- read_data_fileinfoline(8)
dat %>%
  filter(floor_date(Date, "day") == ymd("2018-04-26")) %>%
  head(30)

# File 12, station 68870, 2018-08-21
dat <- read_data_fileinfoline(14)
dat %>%
  filter(floor_date(Date, "day") == ymd("2018-08-21")) %>%
  head(30)
# dat %>%
#   filter(floor_date(Date, "day") == ymd("2018-08-21")) %>%
#   tail(80)

```

### Plot oxygen   
- As mentioned above, 14. Dec 2018 (VR51) seems not to have reached the bottom (or was taken in the wrong place)  
    - Seems likely that bottom oxygen is lower than shown  
- File 3 (plot 5) VR51: Oxygen in Jan and Feb 2019 is very unstable - sensor problem?  
- File 4 (plot 7) VR52: Zero oxygen in March 2017
- File 4 (plot 7) VR52: Zero oxygen in top of water column in Sept 2017 (doesn't matter for classification)  
- File 6 (plot 9) VR52: Error in oxygen in October 2019 (zero)
- File 6 (plot 9) VR52: Error in oxygen in October 2019 (much too high)  
- File 10 (plot 12) 68870: Suspicuous O2 values near bottom in October?
- File 12 (plot 14) 68870: O2 looks a bit suspicuous in Oct 2018, but seems ok
```{r}

plot_timeseries_all("Oksygen")
plot_ctdprofile_all("Oksygen")

```

### Plot oxygen saturation  

```{r}

plot_timeseries_all("Oksygenmetning")
plot_ctdprofile_all("Oksygenmetning")

```
