---
title: "QA for CTDs in Glomfjord - ØKOKYST 2019 report"
author: "DHJ"
date: "13 1 2020"
output:
  html_document:
    keep_md: true
    toc: true
    toc_float: true

---

### QC done January 2020   
- Only CTDs with salinity + temp only    
- See mail from Caroline Mengeot <caroline.mengeot@niva.no> tirsdag 28. januar 2020 10:33  

```{r, echo=FALSE, results='hide', message=FALSE}

## 1. Libraries

library(dplyr)
library(purrr)
library(ggplot2)
library(lubridate)
library(readxl)
library(tidyr)
library(RColorBrewer)

# library(niRvana)

source("12_QA_2019_from_excel_functions.R")
source("13_QA_CTD_2019_functions.R")

# RColorBrewer::display.brewer.all()



```


## 1. Read file   
Only one file in this case, as in script 16 (and in contrast to script 15)  
```{r, echo = FALSE, results = 'hide'}

datafolder <- "K:/Prosjekter/!!! Tiltaksrettet industriovervåkning iht. vannforskriften/16371 - Yara Glomfjorden (tidl.14380)/_2019/Vannsøyle/xlsbase/TilAquamonitor"

# dir(datafolder)
fn <- paste0(datafolder, "/", "glomfjord_CTD_2018-2019_til_Aqm.xlsm")

readxl::excel_sheets(fn)
# Check top of file:
# read_excel(fn, sheet = "Data", n_max = 200)
dat <- read_excel_droplines(fn, sheet = "Data")

# Fix the first column names
# NOT NEEDED

# Add station names
dat_stations <- read_excel(fn, sheet = "Stations")

dat <- dat %>%
  left_join(dat_stations) %>%
  rename(StationCode = StationName)  # for 'plot_ctdprofile_station'

```



## 2. Tables of stations    
```{r, echo = FALSE}
dat %>%
  count(StationCode)

# dat %>%
#   count(SampleDate)
# 
# dat %>%
#   count(StationCode, Date)

stations <- dat %>%
  count(StationCode) %>%
  pull(StationCode)

```


## 3. Salinity  
_NOTE: also see plots for top 20 m further down._  
   
- Gl_5: Profile looks strange at 20-30 m in April 2019 (and no pattern in temperature)  
   
- Very low salinity at 0-1 meter:     
    + Dec 2018 at Gl_3 and Gl_5: Dubious? no indication of top freshwater layer in temperature  
    + 4. April 2019 Gl_5: Dubious? no indication of top freshwater layer in temperature  
    + 21 June 2019 at Gl_6: Dubious? little indication of top freshwater layer in temperature  
    + July 2019 at Gl_2 and Gl_3: probably OK - temperature also indicates top freshwater layer  
    + 15 Aug 2019 at Gl_2 and Gl_4: Dubious? little indication of top freshwater layer in temperature  

```{r, echo = FALSE, fig.width=9, fig.height=7, warning=FALSE}

# Test one station:
# plot_ctdprofile_station("VT16", data = dat, variable = "Saltholdighet", points = FALSE)

stations %>% walk(plot_ctdprofile_station, data = dat, variable = "Saltholdighet", points = FALSE)

```



## 4. Temperature    
_NOTE: also see plots for top 20 m further down._  
  
- Gl_5: 23. April 2019 - Both salinity and temperature looks almost stuck during the top 20 m, but both do in fact vary  
  
```{r, echo = FALSE, fig.width=9, fig.height=7, warning=FALSE}

stations %>% walk(plot_ctdprofile_station, data = dat, variable = "Temperatur", points = FALSE)

```

## 5. Salinity top 20 m 
```{r, echo = FALSE, fig.width=9, fig.height=7, warning=FALSE}

dat20m <- dat %>% filter(Depth1 < 20)

stations %>% walk(plot_ctdprofile_station, data = dat20m, variable = "Saltholdighet", points = FALSE)

```

## 6. Temperature top 20 m 
```{r, echo = FALSE, fig.width=9, fig.height=7, warning=FALSE}

stations %>% walk(plot_ctdprofile_station, data = dat20m, variable = "Temperatur", points = FALSE)

```

## 7. Check Gl_5 at 23. April 2019   
Both salinity and temperature looks almost stuck during the top 20 m   
- but both do in fact vary   
```{r}

dat %>%
   filter(StationCode %in% "Gl_5" & 
            Date > ymd("2019-04-23") & Date < ymd("2019-04-24") &
            Depth1 < 20) %>%
  select(Depth1:Temperatur) %>%
  as.data.frame()



```

