---
title: "QA for CTDs in Norskehavet Sør I - ØKOKYST 2019 report"
author: "DHJ"
date: "13 1 2020"
output:
  html_document:
    keep_md: true
    toc: true
    toc_float: true

---

## QC done January 2021    
- CTDs with salinity + temp + oxygen      
- Excel files from Trond Kristansen's dropbox (29.01.2021)
- Stations:    
    * VT53 - Hardangerfjorden (this script)  
    * VT69 - Hardangerfjorden (this script)   
    * VT70 - Hardangerfjorden (this script)  
    * VT74 - Hardangerfjorden (this script)  
    * VT16 - Sognefjorden (see separate script)   
    * VT79 - Sognefjorden (see separate script)    
    
    ```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}

## 1. Libraries

library(dplyr)
library(purrr)
library(ggplot2)
library(lubridate)
library(readxl)
library(tidyr)
library(RColorBrewer)
library(knitr)
library(kableExtra)

# library(niRvana)

source("12_QA_2019_from_excel_functions.R")
source("13_QA_CTD_2019_functions.R")

# RColorBrewer::display.brewer.all()

knitr::opts_chunk$set(echo = FALSE, results = 'hold')

```


## 1. Read files   

### New data (2020)  
- Read from excel files in folders under `K:/Avdeling/214-Oseanografi/DATABASER/OKOKYST_2017s`  
- 3 files in 3 separate folders in this case    
```{r, echo = FALSE, results = 'hide', message = FALSE}

# datafolder <- "K:/Prosjekter/!!! Tiltaksrettet industriovervåkning iht. vannforskriften/16371 - Yara Glomfjorden (tidl.14380)/_2019/Vannsøyle/xlsbase/TilAquamonitor"


read_excel_files <- FALSE
read_excel_files <- TRUE

if (read_excel_files){
  
  # Files from Trond Kristansen's dropbox (20.02.2021)
  
  datafolder <- "Datasett/Hardangerfjorden_2020"
  fns <- dir(datafolder, pattern = ".xlsx")
  fns <- fns[!grepl("~", fns)]

  read_excel_add_filename <- function(filename, folder){
    fn <- paste0(folder, "/", filename)
    read_excel(fn, na = "-999") %>%
      mutate(Filename = filename)
  }

  dat <- fns %>% map_dfr(read_excel_add_filename, folder = datafolder) %>%
    mutate(Date = as.Date(Date, origin = "1899-12-30")) %>%
    mutate(StationCode2 = paste0(sub("_CTD.xlsx", "", Filename, fixed = TRUE), "_", StationCode),
           StationCode = sub("_CTD.xlsx", "", Filename, fixed = TRUE)
    )
  

  saveRDS(dat, "Data/14_QA_CTD_2020_Hardangerfjorden.rds")
  
} else {
  
  dat <- readRDS("Data/14_QA_CTD_2020_Hardangerfjorden.rds")
  
}

xtabs(~addNA(StationCode), dat)
xtabs(~addNA(Date), dat)

```

#### Compare with old versions

```{r}


if (FALSE){
  
  # install.packages("compareDF")
  library(compareDF)

  # Files from "Old_version" (from Trond 29.01.2021)
  datafolder <- "Datasett/Hardangerfjorden_2020/Old_version/"
  fns <- dir(datafolder, pattern = ".xlsx")
  fns <- fns[!grepl("~", fns)]

  read_excel_add_filename <- function(filename, folder){
    fn <- paste0(folder, "/", filename)
    read_excel(fn, na = "-999") %>%
      mutate(Filename = filename)
  }

  dat_old <- fns %>% map_dfr(read_excel_add_filename, folder = datafolder) %>%
    mutate(Date = as.Date(Date, origin = "1899-12-30")) %>%
    mutate(StationCode2 = paste0(sub("_CTD.xlsx", "", Filename, fixed = TRUE), "_", StationCode),
           StationCode = sub("_CTD.xlsx", "", Filename, fixed = TRUE)
    )
  
  #
  # 1. Compare the whole thing using compareDF::compare_df
  #
  # Should have used Depth1 as grouping as well, but anyway this is too big
  #
  ctab <- compare_df(dat, dat_old, c("StationCode", "Date"))
  
  # Summaries not very informative  
  # ctab$change_count
  ctab$change_summary
  # create_output_table(ctab)  # doesnt make much sense as Depth1 is not included
  
  #
  # 2. Using table() to get an overview on "number of rows" basis
  #
  # Easier to see why there is a difference in no. of records form this:
  #
  tab1 <- dat %>% xtabs(~Date + StationCode, .)
  tab2 <- dat_old %>% xtabs(~Date + StationCode, .)
  tab1 - tab2   
  # - shows that 1 record has been added to VT53 for every date  
  
  #
  # 3. Check one station with *same* number of records   
  #
  # - using compareDF::compare_df plus self-made get_differences_by_variable()
  #
  
  # When using View() for both it is easier to see that the chnages are
  # 1) Row 3 have values for the 4 main sensors in df1, these are NA in version 2
  # 2) Oksygen: changed values
  df1 <-   dat     %>% filter(StationCode %in% "VT69")
  df2 <-   dat_old %>% filter(StationCode %in% "VT69") 
  
  # Group on date and depth - create_output_table gives much better input  
  ctab2 <- compare_df(df1, df2, c("Date", "Depth1"))
  # ctab2$change_count       
  ctab2$change_summary     
  create_output_table(ctab2)  # useful output but very long output
  
  # Easily see that the biggeste differences are for "Oksygen"
  get_differences_by_variable(ctab2)
  
  #
  # Check one station with *different* number of records   
  #
  # - using compareDF::compare_df plus self-made get_differences_by_variable()
  # - get_differences_by_variable() doesn't yet handle different number of rows.,, 
  #
  
  df1 <-   dat     %>% filter(StationCode %in% "VT53" & Date == dat$Date[1]) 
  df2 <-   dat_old %>% filter(StationCode %in% "VT53" & Date == dat$Date[1]) 
  
  # tail(df1)
  # tail(df2)
  
  # As "get_differences_one_variable" is now, we need to do this:
  df1 <- head(df1,-1)   # adjust length so they are equally long
  
  # seems ok:
  sum(df1$Depth1 != df2$Depth1)
  
  # Group on date and depth - create_output_table gives much better input  
  ctab3 <- compare_df(df1, df2, c("Date", "Depth1"))
  ctab3$change_summary     
  
  get_differences_by_variable(ctab3)
  
}


```



### Older data (2017-2019)  

```{r, message = FALSE}
# if (FALSE){

datafolder <- "Datasett/Hardangerfjorden_2019"
fns <- dir(datafolder)

dat_old <- fns %>% map_dfr(read_excel_add_filename, folder = datafolder) %>%
  mutate(Date = as.Date(Date, origin = "1899-12-30")) %>%
  mutate(StationCode2 = paste0(sub("_CTD.xlsx", "", Filename, fixed = TRUE), "_", StationCode),
         StationCode = sub("_CTD.xlsx", "", Filename, fixed = TRUE)
  )


# check dates

cat("-------------------------------------------------------------------------------------- \n")
cat("dat_old: \n")
cat("-------------------------------------------------------------------------------------- \n")
# xtabs(~addNA(Date), dat_old)
xtabs(~addNA(Date) + StationCode, dat_old)

cat("-------------------------------------------------------------------------------------- \n")
cat("dat: \n")
cat("-------------------------------------------------------------------------------------- \n")
# xtabs(~addNA(Date), dat)
xtabs(~addNA(Date) + StationCode, dat)

xtabs(~StationCode2, dat_old)
xtabs(~StationCode2, dat)

#}

```


## 2. Data       

### Check sample dates   
- `dat_old`= 2017-2019   
- `dat` = last part of 2019 + 2020  
```{r}

if (TRUE){
  tab <- xtabs(~Date, dat_old)
  missing <- is.na(dat$Date)
  cat("-------------------------------------------------------------------------------------- \n")
  cat("dat_old: \n")
  cat("-------------------------------------------------------------------------------------- \n")
  cat("Number of dates: ", length(tab), "\n")
  cat("First and last date: ", head(names(tab),1), "-", tail(names(tab),1), "\n")
  cat("Missing dates: ", sum(missing), "\n")
}

tab <- xtabs(~Date, dat)
missing <- is.na(dat$Date)
cat("-------------------------------------------------------------------------------------- \n")
cat("dat: \n")
cat("-------------------------------------------------------------------------------------- \n")
cat("Number of dates: ", length(tab), "\n")
cat("First and last date: ", head(names(tab),1), "-", tail(names(tab),1), "\n")
cat("Missing dates: ", sum(missing), "\n")
  
```


### Dates and max depth of new data    
```{r, results='asis'}

dat %>%
  group_by(StationCode, Date) %>%
  summarize(n = n(), Max_depth = max((Depth1 + Depth2)/2, na.rm = TRUE),
            .groups = "drop") %>%
  arrange(StationCode, Date) %>%
  kbl() %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

# dat %>%
#   count(SampleDate)
# 
# dat %>%2
#   count(StationCode, Date)

stations <- dat %>%
  count(StationCode) %>%
  pull(StationCode  )

```


## 3. Profiles of 2020 data  

### Salinity  
_NOTE: also see plots for top 20 m further down._  
   
- VT69:  
    - Large variation in max. depth  
- VT70:  
    - Dec. 2020: salinity a bit rugged around 180-200 m   

```{r, echo = FALSE, fig.width=9, fig.height=7, warning=FALSE}

# Test one station:
# plot_ctdprofile_station("VT16", data = dat, variable = "Saltholdighet", points = FALSE)

stations %>% walk(
  plot_ctdprofile_station,
  data = dat, variable = "Saltholdighet", points = FALSE)


# stations %>% walk(plot_ctdprofile_station, data = dat, variable = "Saltholdighet", points = FALSE)

```

### Salinity top 50 m 
```{r, echo = FALSE, fig.width=9, fig.height=7, warning=FALSE}

dat50m <- dat %>% filter(Depth1 <= 50)

stations %>% walk(
  plot_ctdprofile_station,
  data = dat50m, variable = "Saltholdighet", points = FALSE)


```


### Temperature    
```{r, echo = FALSE, fig.width=9, fig.height=7, warning=FALSE}

stations %>% walk(
  plot_ctdprofile_station,
  data = dat, variable = "Temperatur", points = FALSE)


```

### Temperature top 50 m 
```{r, echo = FALSE, fig.width=9, fig.height=7, warning=FALSE}

stations %>% walk(
  plot_ctdprofile_station,
  data = dat50m, variable = "Temperatur", points = FALSE)

```

### Oxygen volume      

Somewhat noisy, e.g. April in VT70

```{r, echo = FALSE, fig.width=9, fig.height=7, warning=FALSE}

stations %>% walk(
  plot_ctdprofile_station,
  data = dat, variable = "Oksygen", points = FALSE)

```

### Oxygen saturation        
```{r, echo = FALSE, fig.width=9, fig.height=7, warning=FALSE}

stations %>% walk(
  plot_ctdprofile_station,
  data = dat, variable = "Oksygenmetning", points = FALSE)


```




## 4. Times series (since 2017)  

### Combined data
```{r}

dat2 <- bind_rows(dat_old, dat)

# xtabs(~addNA(Date), dat2)

dat2 %>%
  distinct(Date, StationCode) %>%
  mutate(Year = year(Date)) %>%
  xtabs(~Year + addNA(StationCode), .)

# dat3b

```

### All stations  
```{r}

  
  # df <- df_chem %>%
  #   distinct(STATION_CODE, PARAM_NAME) 
  # 
  # walk2(
  #   df$STATION_CODE[1:3],
  #   df$PARAM_NAME[1:3],
  #   ~plot_ctdprofile_station(stationcode = .x, 
  #                            variable = .y,
  #                            data = dat, 
  #                            points = TRUE, year_by_month = TRUE)
  


```

### Station VT53       
```{r}

if (FALSE){
  # check depths
  dat2 %>%
    filter(StationCode %in% "VT53") %>%
    group_by(Date) %>%
    summarize(Max_depth = max(Depth1)) %>%
    arrange(Max_depth)
}

# debugonce(plot_timeseries_station)
plot_timeseries_station(stationcode = "VT53", data = dat2, variable = "Saltholdighet", titletext = "")
plot_timeseries_station(stationcode = "VT53", data = dat2, variable = "Temperatur", titletext = "")
plot_timeseries_station(stationcode = "VT53", data = dat2, variable = "Oksygen", titletext = "")
plot_timeseries_station(stationcode = "VT53", data = dat2, variable = "Oksygenmetning", titletext = "")

```


### Station VT69       
```{r}

if (FALSE){
  # check depths
  dat2 %>%
    filter(StationCode %in% "VT69") %>%
    group_by(Date) %>%
    summarize(Max_depth = max(Depth1)) %>%
    arrange(Max_depth)
}

# debugonce(plot_timeseries_station)
plot_timeseries_station(stationcode = "VT69", data = dat2, variable = "Saltholdighet", titletext = "")
plot_timeseries_station(stationcode = "VT69", data = dat2, variable = "Temperatur", titletext = "")
plot_timeseries_station(stationcode = "VT69", data = dat2, variable = "Oksygen", titletext = "")
plot_timeseries_station(stationcode = "VT69", data = dat2, variable = "Oksygenmetning", titletext = "")

```

### Station VT70       
```{r}

if (FALSE){
  # check depths
  dat2 %>%
    filter(StationCode %in% "VT70") %>%
    group_by(Date) %>%
    summarize(Max_depth = max(Depth1)) %>%
    arrange(Max_depth)
}

# debugonce(plot_timeseries_station)
plot_timeseries_station(stationcode = "VT70", data = dat2, variable = "Saltholdighet", titletext = "")
plot_timeseries_station(stationcode = "VT70", data = dat2, variable = "Temperatur", titletext = "")
plot_timeseries_station(stationcode = "VT70", data = dat2, variable = "Oksygen", titletext = "")
plot_timeseries_station(stationcode = "VT70", data = dat2, variable = "Oksygenmetning", titletext = "")

```

### Station VT74       
```{r}

if (FALSE){
  # check depths
  dat2 %>%
    filter(StationCode %in% "VT74") %>%
    group_by(Date) %>%
    summarize(Max_depth = max(Depth1)) %>%
    arrange(Max_depth)
}

# debugonce(plot_timeseries_station)
plot_timeseries_station(stationcode = "VT74", data = dat2, variable = "Saltholdighet", titletext = "")
plot_timeseries_station(stationcode = "VT74", data = dat2, variable = "Temperatur", titletext = "")
plot_timeseries_station(stationcode = "VT74", data = dat2, variable = "Oksygen", titletext = "")
plot_timeseries_station(stationcode = "VT74", data = dat2, variable = "Oksygenmetning", titletext = "")

```



