---
title: "ØKOKYST Ferrybox"
author: "Helene"
date: "7 februar 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## 1. Libraries
```{r}

library(plyr)      # used by functions in "Get_files_NIVA_ftp_server_functions.R" 
library(dplyr)     # load AFTER plyr
library(purrr)
library(ggplot2)
library(lubridate)
library(readxl)
library(tidyr)
library(RColorBrewer)

source("Get_files_NIVA_ftp_server_functions.R")

# library(niRvana)
#source("12_QA_2019_from_excel_functions.R")
# RColorBrewer::display.brewer.all()
```

## 2.Aquamonitor functions
-code from Dag to read and reformat Aquamonitor data (email 07.02.2020). NB! columns must start with stationId and and with Depth2
```{r}
# READING "WIDE" AQUAMONITOR DATA ----
#

AqMexport_read_waterchemistry <- function(filename, reformat_long = TRUE, remove_duplicates = TRUE, sheetname = "WaterChemistry"){
  # Read top line, only for the column names
  df_names <- read_excel(fn, sheet = sheetname, col_names = TRUE, n_max = 0)
  # Read data. All data is read as text (strings) and later converted, due to the "<" signs
  df_chem <- read_excel(fn, sheet = "WaterChemistry", col_names = FALSE, skip = 2, col_types = "text")
  # Set column names (copy them from df_names)
  if (ncol(df_names) == ncol(df_chem)){
    names(df_chem) <- names(df_names)
  } else {
    cat("Not same number of columns!\n")
  }
  # Convert these variables to numeric
  for (col in c("ProjectId", "SampleDate", "Time", "Depth1", "Depth2")){
    df_chem[[col]] <- as.numeric(df_chem[[col]])
  }
  # UNIX time
  # We just overwrite Time with UNIX time, which is usually just 00:00 anyway
  df_chem$Time <- as.POSIXct((df_chem$SampleDate - 25569)*24*3600, origin = "1970-01-01", tz = "GMT")
  
  # Reformat data to long/narrow format (default option)
  if (reformat_long)
    df_chem <- AqMexport_reformat_long(df_chem, remove_duplicates = remove_duplicates)
  
  df_chem 
  }

# Note hard-coded columns "ProjectId:Depth2"
AqMexport_reformat_long <- function(dat, remove_duplicates = TRUE){
  dat_long <- dat %>%
    pivot_longer(cols = -c(ProjectId:Depth2), names_to = "Variable", values_to = "Value_chr") 
  # Get numeric data values
  x <- sub(",", ".", dat_long$Value_chr, fixed = TRUE)
  x <- sub(",", ".", x, fixed = TRUE)
  dat_long$Value <- as.numeric(sub("<", "", x))
  # Make less-than flag
  dat_long$Flag <- ifelse(grepl("<", dat_long$Value_chr, fixed = TRUE), "<", NA)
  # Get rid of duplicates (several observations, as there are several projects and each value isrepeated for each project)
  if (remove_duplicates)
    dat_long <- AqMexport_remove_duplicates(dat_long)
  dat_long
}
  
AqMexport_remove_duplicates <- function(dat){
  dat %>%
    filter(!is.na(Value)) %>%
    group_by(StationId, StationCode, StationName, Time, Depth1, Depth2, Variable) %>%
    summarise(ProjectId = paste(ProjectId, collapse = ","),
              ProjectName = paste(ProjectName, collapse = "; "),
              Value = first(Value), 
              Flag = first(Flag)) %>%
    ungroup()
}

```

### test using Aqm read functions
*NB fikk kke til å fungere
```{r}

# dat_broad <- AqMexport_read_waterchemistry("Datasett/AqM_2017_2019_ØKOKYST_Ferrybox.xlsx", reformat_long = FALSE)   # 30 seconds
# dat1 <- AqMexport_reformat_long(dat_broad, remove_duplicates = FALSE)
# dat2 <- AqMexport_remove_duplicates(dat1)

```


## 2. Read file
```{r}

# DOC	  Fluo	  KlfA	NH4-N	NO3 	PO4-P	  Salt	SiO2	Temperatur	TOC	TOTN	TOTP	TSM	Turbiditet
# mg/L C	µg/l	µg/l	µg/l	µg/l	µg/l		      mg/l	C	          mg/l µg/l	µg/l	mg/l	FNU

# sjekke filer i folder "Datasett"
#dir("Datasett")

df_aqm <- read_excel("Datasett/AqM_2017_2019_ØKOKYST_Ferrybox_ToR.xlsx", sheet = "WaterChemistry")

#lage datoobjekt
df_aqm$Date <- as.Date (paste (df_aqm$Year, df_aqm$Month, df_aqm$Day, sep ="-"))

str(df_aqm)
summary(df_aqm)


# no duplicates
str (test <- distinct (df_aqm))

#include only 2018 and 2019 data
df_aqm <- subset(df_aqm, Date >= "2018-01-01" &  Date <= "2019-12-31")

# StationCode seems to be correct for all rows
#levels(as.factor (df_aqm$StationName))
#levels(as.factor (df_aqm$StationCode))
df_aqm$StationCode <- as.factor (df_aqm$StationCode)
df_aqm$StationName <- as.factor (df_aqm$StationName)
levels(df_aqm$StationCode)

#order stations from south to north
df_aqm$StationCode <- factor(df_aqm$StationCode, levels = c("VT4", "VT12", "VT72", "VT80", "VT45", "VT22", "VT23", "VT76", "VR23", "VR25"))

# convert some columns read as chr to numeric
df_aqm<- df_aqm %>% mutate(
    KlfA = as.numeric(KlfA), NH4 = as.numeric(NH4), NO3_NO2 = as.numeric(NO3_NO2),  PO4 = as.numeric(PO4) 
)

# rounding
# df_aqm <- df_aqm %>% 
#   mutate_if(is.numeric, round, digits = 2)

# look at missing values
tb <- df_aqm %>% 
  gather("Variable", Value, DOC:TOTP) %>%
  filter(!is.na(Value)) %>%
  xtabs(~StationCode + Variable, .)
tb

```

### Converting DOC from mg C /l (NIVAlab) to umol C /Kg
```{r}
# https://ocean.ices.dk/Tools/UnitConversion.aspx
#Conversions
#??g/l = ??mol/l × MW
#g/l ??? g/kg × 1.025

# converting mg/l to ug/L
DOC_ug <- df_aqm$DOC * 1000

# converting ug/l to umol/L
df_aqm$DOC_umolL <- DOC_ug / 12.01

# converting umol/l to umol/kg
df_aqm$DOC_umolKg <- df_aqm$DOC_umolL / 1.025



```


## 3. Plot DOC across FB stations
```{r}

#, scales = "free_y"
df_aqm %>%
  ggplot(aes(Date, DOC)) + 
    geom_point() + geom_line() +
    facet_wrap(~StationCode, ncol=2)

#ggsave ("Figurer/DOC_FBstations.png", width = 8, height = 6, dpi=500)

# plot DOC with label every 2nd months
df_aqm %>%
  ggplot(aes(Date, DOC)) + 
    geom_point() + geom_line() +
    facet_wrap(~StationCode, ncol=2) +
      labs(title = "", x = "", y = "DOC (mg/L)") +
        scale_x_date(date_breaks = "2 month", date_labels =  "%m-%y") +
        theme(axis.text.x=element_text(angle=60, hjust=1))

ggsave ("Figurer/DOC_FBstations_V2.png", width = 8, height = 6, dpi=500)

```

## 4. Plot allvariables across FB stations
```{r}


# plot Chla
df_aqm %>%
  ggplot(aes(Date, KlfA)) + 
    geom_point() + geom_line() +
    facet_wrap(~StationCode, ncol=2) +
      labs(title = "", x = "", y = "Chla (µg/l)") +
        scale_x_date(date_breaks = "2 month", date_labels =  "%m-%y") +
        theme(axis.text.x=element_text(angle=60, hjust=1))

ggsave ("Figurer/Chla_FBstations_V2.png", width = 8, height = 6, dpi=500)

# plot Chla
df_aqm %>%
  ggplot(aes(Date, KlfA)) + 
    geom_point() + geom_line() +
    facet_wrap(~StationCode, ncol=2) +
      labs(title = "", x = "", y = "Chla (µg/l)") +
        scale_x_date(date_breaks = "2 month", date_labels =  "%m-%y") +
        theme(axis.text.x=element_text(angle=60, hjust=1))

ggsave ("Figurer/Chla_FBstations_V2.png", width = 8, height = 6, dpi=500)

# plot No3+NO2
df_aqm %>%
  ggplot(aes(Date, NO3_NO2)) + 
    geom_point() + geom_line() +
    facet_wrap(~StationCode, ncol=2) +
      labs(title = "", x = "", y = "NO3 (µg/l)") +
        scale_x_date(date_breaks = "2 month", date_labels =  "%m-%y") +
        theme(axis.text.x=element_text(angle=60, hjust=1))

ggsave ("Figurer/NO3_FBstations_V2.png", width = 8, height = 6, dpi=500)

# plot NH4
df_aqm %>%
  ggplot(aes(Date, NH4)) + 
    geom_point() + geom_line() +
    facet_wrap(~StationCode, ncol=2) +
      labs(title = "", x = "", y = "NH4 (µg/l)") +
        scale_x_date(date_breaks = "2 month", date_labels =  "%m-%y") +
        theme(axis.text.x=element_text(angle=60, hjust=1))

ggsave ("Figurer/NH4_FBstations_V2.png", width = 8, height = 6, dpi=500)

# plot PO4
df_aqm %>%
  ggplot(aes(Date, PO4)) + 
    geom_point() + geom_line() +
    facet_wrap(~StationCode, ncol=2) +
      labs(title = "", x = "", y = "PO4 (µg/l)") +
        scale_x_date(date_breaks = "2 month", date_labels =  "%m-%y") +
        theme(axis.text.x=element_text(angle=60, hjust=1))

ggsave ("Figurer/PO4_FBstations_V2.png", width = 8, height = 6, dpi=500)

# plot TotN
df_aqm %>%
  ggplot(aes(Date, TOTN)) + 
    geom_point() + geom_line() +
    facet_wrap(~StationCode, ncol=2) +
      labs(title = "", x = "", y = "TOTN (µg/l)") +
        scale_x_date(date_breaks = "2 month", date_labels =  "%m-%y") +
        theme(axis.text.x=element_text(angle=60, hjust=1))

ggsave ("Figurer/TOTN_FBstations_V2.png", width = 8, height = 6, dpi=500)

# plot TotP
df_aqm %>%
  ggplot(aes(Date, TOTP)) + 
    geom_point() + geom_line() +
    facet_wrap(~StationCode, ncol=2) +
      labs(title = "", x = "", y = "TOTP (µg/l)") +
        scale_x_date(date_breaks = "2 month", date_labels =  "%m-%y") +
        theme(axis.text.x=element_text(angle=60, hjust=1))

ggsave ("Figurer/TOTP_FBstations_V2.png", width = 8, height = 6, dpi=500)

```

## 5a.  Ferrybox log data 2019
We do this before 2018 in order to get the headers (column names),
which we will use for 2018 data    
- Data downloaded from Pierre's OneDrive 12.02.2020    
- [Link to Pierre's OneDrive](https://niva365-my.sharepoint.com/:f:/g/personal/pierre_jaccard_niva_no/Et9g2CZL2_tEg3vOt7ACQh0BHCjrOcKzPz-iAgdYst2ddQ?e=hpkNeW)  

```{r}

df_ferrybox_sampledata_2019_fa <- readxl::read_excel(
  "Datasett/Ferrybox_samples_OneDrive/FA_2019_automatic_samples.xlsx")

df_ferrybox_sampledata_2019_tf <- readxl::read_excel(
  "Datasett/Ferrybox_samples_OneDrive/TF_2019_automatic_samples.xlsx")

# names(df_ferrybox_sampledata_2019_fa)
# names(df_ferrybox_sampledata_2019_tf)

X <- df_samples_fa$SYSTEM_DATE_DMY - ymd_hms("1899-12-31 00:00:00")
df_samples_fa <- df_samples_fa %>%
  mutate(TIME = SYSTEM_TIME + X)

# Check
# df_samples_fa$SYSTEM_DATE_DMY %>% head()
# df_samples_fa$SYSTEM_TIME %>% head()
# df_samples_fa$TIME %>% head()

unique(df_samples_fa$SYSTEM_DATE_DMY)

```

## 5b. Ferrybox log data 2018   
Get dates and filenames ('df_logfilenames')  
```{r}

get_folders()

#
# Fantasy
#

# Get the url for the vessel we want
url.folder <- get_url_foldername("fantasy")

# Get file names of all log files for that vessel
samplefiles_fn <- get_filenames_samplefiles("fantasy")

head(samplefiles_fn)
# Check
# tail(logfiles_fn)

# Combine with dates 
df_samplefilenames_fa_2018 <- data.frame(
  filename = samplefiles_fn,
  Date = ymd(substr(samplefiles_fn, 9, 16)),   # note that 9,16 is specifically for 'samples'
  stringsAsFactors = FALSE) %>%
  filter(year(Date) == 2018)


#
# Trollfjord
#

# Get the url for the vessel we want
url.folder <- get_url_foldername("trollfjord")

# Get file names of all log files for that vessel
samplefiles_fn <- get_filenames_samplefiles("trollfjord")

# Check
# tail(logfiles_fn)

# Combine with dates 
df_samplefilenames_tf_2018 <- data.frame(
  filename = samplefiles_fn,
  Date = ymd(substr(samplefiles_fn, 9, 16)), # note that 9,16 is specifically for 'samples'
  stringsAsFactors = FALSE) %>%
  filter(year(Date) == 2018)

# Combine
df_samplefilenames_2018 <- 
  bind_rows(
    df_samplefilenames_fa_2018 %>% mutate(Ferrybox = "fantasy"),
    df_samplefilenames_tf_2018 %>% mutate(Ferrybox = "trollfjord")
  )

```

### Download all  
Or just use saved log data
```{r}

redownload_data <- FALSE  # if FALSE, we just read the saved data


if (redownload_data){  # takes just a couple of minutes
  
  df_ferrybox_sampledata_2018_fa <- get_several_logfiles(
    df_samplefilenames_fa_2018$filename, 
    vessel = "fantasy",
    trace = FALSE)
  
  df_ferrybox_sampledata_2018_tf <- get_several_logfiles(
    df_samplefilenames_tf_2018$filename, 
    vessel = "trollfjord",
    trace = FALSE)
  
  names(df_ferrybox_sampledata_2018_fa) <- names(df_ferrybox_sampledata_2019_fa)
  names(df_ferrybox_sampledata_2018_tf) <- names(df_ferrybox_sampledata_2019_tf)

  saveRDS(df_ferrybox_sampledata_2018_fa,
          "Datasett/18_df_ferrybox_sampledata_2018_fa.rds")
  saveRDS(df_ferrybox_sampledata_2018_tf,
          "Datasett/18_df_ferrybox_sampledata_2018_tf.rds")
  
  
} else {
  
  df_ferrybox_sampledata_2018_fa <-
    readRDS("Datasett/18_df_ferrybox_sampledata_2018_fa.rds")
  df_ferrybox_sampledata_2018_tf <-
    readRDS("Datasett/18_df_ferrybox_sampledata_2018_tf.rds")
  
}


```


### Combine years  
```{r}
#
# FA
#
df_ferrybox_sampledata_2018_fa[1:7] %>% str()
df_ferrybox_sampledata_2019_fa[1:7] %>% str()

df_ferrybox_sampledata_2018_fa <- df_ferrybox_sampledata_2018_fa %>%
  mutate(SYSTEM_DATE_DMY = dmy_hms(paste(SYSTEM_DATE_DMY, "00:00:00")),
         SYSTEM_TIME = ymd_hms(paste("1899-12-31", SYSTEM_TIME)))

df_ferrybox_sampledata_fa <- bind_rows(
  df_ferrybox_sampledata_2018_fa,
  df_ferrybox_sampledata_2019_fa)

#
# Trollfjord
#
df_ferrybox_sampledata_2018_tf[1:7] %>% str()
df_ferrybox_sampledata_2019_tf[1:7] %>% str()

df_ferrybox_sampledata_2018_tf <- df_ferrybox_sampledata_2018_tf %>%
  mutate(SYSTEM_DATE = dmy_hms(paste(SYSTEM_DATE, "00:00:00")),
         SYSTEM_TIME = ymd_hms(paste("1899-12-31", SYSTEM_TIME)))

df_ferrybox_sampledata_tf <- bind_rows(
  df_ferrybox_sampledata_2018_tf,
  df_ferrybox_sampledata_2019_tf)

names(df_ferrybox_sampledata_fa)
names(df_ferrybox_sampledata_tf)

# df_ferrybox_sampledata <- 
#   bind_rows()



```






















## 6. Station positions and dates   
df_station_date
```{r}

df_station_pos <- read_excel("Datasett/AqM_2017_2019_ØKOKYST_Ferrybox_ToR.xlsx", sheet = "StationPoint")

my_map <- map_data("world", c("Norway", "Sweden"))

ggplot(df_station_pos, aes(x = Longitude, y = Latitude)) +
  geom_text(aes(label = StationCode)) +
  annotation_map(my_map, fill = NA, color = "blue")

df_station_pos <- df_station_pos %>%
  mutate(Ferrybox = ifelse(StationCode %in% "VT4", "fantasy", "trollfjord"))

# One line per station/date combination in data
df_station_date <- df_aqm %>%
  count(StationCode, Date) %>%
  left_join(subset(df_station_pos, select = c(StationCode, Latitude, Longitude, Ferrybox)))

df_station_date

```

### Check sample files
```{r}

df <- read.table("Datasett/samples_20180314-083000.txt")
df <- read.table("Datasett/samples_20180420-081001.txt")
names(df) <- colnames.TF
names(df)[29:33] <- paste0("X", 1:5)
  
library(leaflet)
leaflet() %>%
  addTiles() %>%
  addMarkers(lng = df$Lon, lat = df$Lat,
             popup = df$ASAMP)

```



## 7. Get list of ferrybox data files to download  

### Get closest Ferrybox date for all lab-data dates  
'df_logfilenames_closest'  
```{r}

# Functrion for getting Ferrybox dat closest to the given 'date'
# Returns a data frame
get_closest_day <- function(date, ferryboxline, ferryboxfilenames){
  ferryboxfilenames %>%
    filter(Ferrybox %in% ferryboxline) %>%
    mutate(Days_diff = (Date - date)/ddays(1)) %>%                     # difference in number of days
    filter(abs(Days_diff) %in% min(abs(Days_diff), na.rm = TRUE)) %>%  # get the smallest difference
    rename(Date_ferrybox = Date) %>%  # To avoid confusion
    mutate(Date_lab = date, Ferrybox = ferryboxline)
    }
# Test

get_closest_day(ymd("2018-07-20"), "fantasy", ferryboxfilenames = df_logfilenames)
get_closest_day(ymd("2018-07-20"), "trollfjord", ferryboxfilenames = df_logfilenames)

get_closest_day(ymd("2018-07-20"), data = df_tf_logfilenames)

lab_dates <- df_station_date %>%
  count(Ferrybox, Date)
    
lab_dates <- unique(df_station_date$Date)
names(lab_dates) <- lab_dates  # Makes it possible to include lab_dates in the final result (using '.id = ...')   

df_logfilenames_closest_day <- lab_dates %>% 
  map_dfr(get_closest_day, data = df_fa_logfilenames, .id = "Date_labdata") %>%
  mutate(Date_labdata = ymd(Date_labdata))
 <- dates %>% map_dfr(get_closest, .id = "Date_labdata")

```

### Add ferrybox data file name to 'df_station_date'  
```{r}

df_station_date <- df_station_date %>%
  left_join(df_logfilenames_closest, by = c("Date" = "Date_labdata"))

```

### Check lab dates that were not found in Ferrybox  
```{r}

dfx <- df_station_date %>%
  filter(Days_diff != 0) %>%
  select(StationCode, Date, Date_ferrybox, Days_diff) %>%
  arrange(Date)

dfx

write.table(dfx, "clipboard", sep = "\t")

```

## 8. Get all ferrybox data  
A couple of minutes  
```{r}
fb_data <- get_several_logfiles(
  unique(df_station_date$filename), 
  vessel = "fantasy",
  trace = FALSE)

#dim(fb_data)
#test <- head(fb_data)
#colnames(test) <- colnames.FA
#test 

colnames(fb_data) <- colnames.FA

names(fb_data)[32:33] <- c("X1", "X2")

# 
# Time
#
fb_data <- fb_data %>%
  rename(Date_text = Date) %>%
  mutate(Date = dmy(Date_text),
         Time = dmy_hms(paste(Date_text, TimeOfDay)),
         Month = month(Time),
         Year = year(Time)
  )



```

## 9. Select ferrybox data for each position  

### Funcion: find closest latitude  
```{r}
# sort(unique(fb_data$Date)) %>% head()

# Hard-coded: variable names 'Date', 'Lat'
# Returns data frame
getdata_closest_latitude <- function(latitude, data, max_difference = 0.1){
  df <- data %>%
    group_by(Date) %>%
    mutate(Lat_diff = Lat - latitude) %>%
    filter(abs(Lat_diff) == min(abs(Lat_diff), na.rm = TRUE)) %>%
    filter(Lat_diff <= max_difference) %>%
    arrange(Date)
  check <- df %>%
    count(Date)
  cat("Lat:", latitude, "returned maximum", max(check$n, na.rm = TRUE), "lines per date\n")
  df
}
# Test 1
# df <- get_closest_latitude(57, fb_data) #  %>% select(Time, Lat, Lon, Lat_diff, TEMP_IN)

# Test 2
getdata_closest_latitude(59.7, fb_data) %>%
  ggplot(aes(Date, TEMP_IN)) +
  geom_line() + geom_point()
    
```

### Select ferrybox data for each station   
Here, we use latitude only as the route is pretty much N-s    
```{r}
# LAtitudes to use to select data
latitudes <- df_station_pos$Latitude
names(latitudes) <- df_station_pos$StationCode

# Select all these 
fb_data_selected <- latitudes %>% map_dfr(getdata_closest_latitude, 
                                          data = fb_data, 
                                          .id = "StationCode")
fb_data_selected <- fb_data_selected %>%
  rename(Date_ferrybox = Date)

fb_data_selected %>%
  count()

```

### Add ferrybox data to df_aqm  
```{r}

nrow(df_aqm)
df_aqm <- df_aqm %>%
  left_join(subset(df_logfilenames_closest, 
                   select = c(Date_labdata, Date_ferrybox, Days_diff)),
            by = c("Date" = "Date_labdata")) %>%
  left_join(subset(fb_data_selected, 
                   select = c(StationCode, Date_ferrybox, TimeOfDay, Lon, Lat, 
                              PUMP, VAL, TURB, FLU, TEMP_IN, SAL, OX_CONC)),
            by = c("StationCode", "Date_ferrybox"))
nrow(df_aqm)

```


```{r}

ggplot(df_aqm, aes(Date, TEMP_IN, color = StationCode)) +
  geom_line() + geom_point()

```


```{r}
# Dates in data
i <- 1
pos <- c(df_station_pos$Longitude[i], df_station_pos$Latitude[i])

get_closest_pos <- function(pos)

fb_data %>%
  arrange(Date)
  
  
df_dates <- unique(df$Date)

plot()

#
# Read latest 20 files
#
# debugonce(get_several_logfiles )
fb_data <- get_several_logfiles(tail(df_files_fn, 20), vessel = "fantasy")

#dim(fb_data)
#test <- head(fb_data)
#colnames(test) <- colnames.FA
#test 

colnames(fb_data) <- colnames.FA

# 
# Time
#
fb_data$Time <- dmy(fb_data$Date) + hms(fb_data$TimeOfDay)
fb_data$Date <- floor_date(fb_data$Time, "day")
fb_data$Month <- month(fb_data$Time)

```

