---
title: "12_QA_2019_from_excel"
author: "DHJ"
date: "13 1 2020"
output: 
  html_document:
    keep_md: true
    toc: true
    toc_float: true
---

# QA for Nordsjøen Nord, nutrients, ØKOKYST 2019 report (January 2020)    


  
## 1. Libraries
```{r, echo = FALSE, results = 'hide', message = FALSE}

library(dplyr)
library(purrr)
library(ggplot2)
library(lubridate)
library(readxl)
library(tidyr)
library(RColorBrewer)

# library(niRvana)

source("12_QA_2019_from_excel_functions.R")
source("13_QA_CTD_2019_functions.R")

# RColorBrewer::display.brewer.all()
```


## 2. Read file    
```{r, echo = FALSE, results = 'hide'}

fn <- "Datasett/Nordsjø Nord/QC_hydrokjemi_NordsjøenNord_2017_2019.xlsx"
readxl::excel_sheets(fn)
dat <- read_excel_droplines(fn, sheet = "Sheet1")

# Fix the first column names
dat_row2 <- read_excel(fn, sheet = "Sheet1", skip = 1, n_max = 1)
names(dat)[1:8] <- names(dat_row2)[1:8]

# Fixing 1 
dat <- dat %>%
  rename(Saltholdighet = Salinitet)  %>%   # for plot_ctdprofile_station to work
  mutate(Date = dmy_hms(SampleDate))

# Fixing 2 - change variablse to numeric  
vars <- names(dat %>% select(KlfA:TSM))
for (var in vars)
  dat[[var]] <- as.numeric(dat[[var]])

```

### Tables of stations    
```{r, echo = FALSE}
dat %>%
  count(StationCode)

# dat %>%
#   count(SampleDate)
# 
# dat %>%
#   count(StationCode, Date)

stations <- dat %>%
  count(StationCode) %>%
  pull(StationCode)

```


```{r variables, echo = FALSE, results = 'hide'}
vars
```


## 3. KlfA
```{r, echo = FALSE, fig.width=9, fig.height=7, warning=FALSE}

# debugonce(plot_ctdprofile_station)
# plot_ctdprofile_station("VT16", data = dat, variable = "KlfA", points = TRUE)

stations %>% walk(plot_ctdprofile_station, data = dat, variable = "KlfA", points = TRUE)

```


## 4. SiO2  
- Some really high values (ca 3 or more) in   
    + VT52 (Jan 2018 and Aug 2018)  
    + VT53 (Jan 2018)  
```{r, echo = FALSE, fig.width=9, fig.height=7, warning=FALSE}

stations %>% walk(plot_ctdprofile_station, data = dat, variable = "SiO2", points = TRUE)

```


## 5. Nitrogen    
  
### Nitrite + nitrate, ammonium, Tot-N  
- TOTN is too low (>20%, see next section) in  
    + VT52, July 2018, 20 m  
    + VT74, Aug 2018, 30 m  
    + VT75, Dec 2017, 10 m  
- TOTN seems too low (deviates from pattern in inorganic components) in   
    + VT52, Aug 2018, 0, 5 and 10 m
    + VT53, Jan 2018, 5 m
    + VT53, July 2018, 5 + 10 m
    + VT70, July 2018?
    + VT70, Dec 2018, 5 m
    + VT74, May 2018, 0 m
    + VT75, Dec 2017, 10 m
- TOTN seems too high in   
    + VT75, Oct 2018, 10 m
```{r, echo = FALSE, fig.width=9, fig.height=7, warning=FALSE}

dat <- dat %>%
  mutate(Inorg_N_sum = `NH4-N` + `NO3+NO2-N`)

stations %>% walk(plot_ctdprofile_station_multi, 
                  data = dat, variables = c("NH4-N", "NO3+NO2-N", "Inorg_N_sum", "TOTN"), 
                  points = TRUE, titletext2 = "Nitrogen")

```


### Inorg. N as fraction of TOTN  
```{r, echo = FALSE, fig.width=9, fig.height=7, warning=FALSE}

dat <- dat %>%
  mutate(Inorg_N_fraction = Inorg_N_sum/TOTN)

dat %>%
  filter(Inorg_N_fraction > 1) %>%
  group_by(StationCode, Date) %>%
  summarize(N = n(), 
            Inorg_N_fraction_max = max(Inorg_N_fraction),
            Inorg_N_fraction_mean = mean(Inorg_N_fraction)
            )

stations %>% walk(plot_ctdprofile_station, data = dat, variable = "Inorg_N_fraction", points = TRUE,
                  limits = 1)


```

## 6. Phosphorus

### PO4-P and TOTP   
- TOTP is suspiciously high:  
    + VT70, Sept 2017, 20 m
    + VT70, Aug 2019, 5 m
- TOTP is a little too low (5-13%, see next section):
    + VT16, March and May 2017  
- PO4 is suspiciously low:  
    + VT74, April 2017, 30 m  
    
```{r, echo = FALSE, fig.width=9, fig.height=7, warning=FALSE}

# debugonce(plot_ctdprofile_station_multi)
# plot_ctdprofile_station_multi("VT16", data = dat, variables = c("PO4-P", "TOTP"), 
#                               points = TRUE, titletext2 = "Phosporus")

stations %>% walk(plot_ctdprofile_station_multi, 
                  data = dat, variables = c("PO4-P", "TOTP"), 
                  points = TRUE, titletext2 = "Phosporus")

```




### Inorg. P as fraction of TOTP
```{r, echo = FALSE, fig.width=9, fig.height=7, warning=FALSE}

dat <- dat %>%
  mutate(Inorg_P_fraction = `PO4-P`/TOTP)

dat %>%
  filter(Inorg_P_fraction > 1) %>%
  group_by(StationCode, Date) %>%
  summarize(N = n(), 
            Inorg_P_fraction_max = max(Inorg_P_fraction),
            Inorg_P_fraction_mean = mean(Inorg_P_fraction)
            )

stations %>% walk(plot_ctdprofile_station, data = dat, variable = "Inorg_P_fraction", points = TRUE,
                  limits = 1)


```






