---
title: "17. QC Norskehavet Nord I"
author: "DHJ"
date: "16 2 2020"
output: 
  html_document:
    toc: true
    toc_float: true
    keep_md: true
---

**QC of Norskehavet Sør I (Caroline + Lars)**      
- Checked input to these files:  
`K:\Prosjekter\Sjøvann\KYSTOVERVÅKING ØKOKYST\KYSTOVERVÅKING ØKOKYST 2017-2020\ØKOKYST DP Norskehavet Sør I O-17091_18091_19091_200091\Rapport 2019-data\klassification`
- Stations:
    + VT71 Skinnbrokleia
    + VR51 Korsen
    + VT72 Herøyfjorden (Ferrybox station)


   
**Veileder:**  
- `Documents\artikler\veiledere etc\02_2013_klassifiserings-veileder_.pdf`  
- Chlorophyll: page 91.   
"I SørNorge (til Stadt) anbefales det at innsamlingen starter i februar og avsluttes ved utgangen av oktober."   
"Nord for Stadt anbefales det at innsamlingsperioden strekker seg fra mars til og med september."   
90% percentiles should be used   
  
- Also see calculation example on on page 187-189, and the comment on page 189:  
"Det gjøres oppmerksom på at dataprogrammene som beregner 90-persentil, gjerne benytter ulike metoder". We use the Excel method PERCENTILE.EXC, corresponding to type 6 quantiles in R  
  
- Nutrients: page 102 in veileder  
Vinter er november til og med februar   
Sommer er juni til og med august  

  
## 0. Libraries   
If you ony want to look at Norskehavet Sør 1  you can skip to 11 after this chunk      
```{r}

library(dplyr)
library(purrr)
library(ggplot2)
library(lubridate)
library(readxl)
library(tidyr)
library(knitr)         
library(RColorBrewer)

# library(niRvana)

source("12_QA_2019_from_excel_functions.R")
source("13_QA_CTD_2019_functions.R")          # used in section 15  


# RColorBrewer::display.brewer.all()
```



## 1. Read Norskehavet Nord I files   
Used by Caroline     
```{r}
# Based on section 5 above

fn <- "K:/Avdeling/214-Oseanografi/DATABASER/OKOKYST_2017/AqM_download_2019/AqM_2019_NH_Sor1_des2019.xlsx"

dat_broad <- AqMexport_read_waterchemistry(fn, reformat_long = FALSE)   # 30 seconds
dat1 <- AqMexport_reformat_long(dat_broad, remove_duplicates = FALSE)
dat2 <- AqMexport_remove_duplicates(dat1)

xtabs(~StationCode + ProjectId, dat2)

# Skinnbrokleia = VT71!
# delete duplicate data in ProjectID "6342,11946", "6342,6342,11946,11946" 
dat2 <- dat2 %>%
  filter(StationCode != "Skinnbrokleia")   

```

### Checks
```{r}

# Check
xtabs(~Variable, dat2)
xtabs(~Variable + StationCode, dat2)

# Check - note that 0.5 m was used instead of 0 m before 2017  
dat2 %>%
  filter(Variable %in% "NO3+NO2-N") %>%
  xtabs(~Depth1 + year(Time), .)

# Check
# NO3+NO2-N and NH4-N used all years  
# NO2 + NO3 not used
dat2 %>%
  filter(Variable %in% c("TOTN", "NH4-N", "NO3+NO2-N", "NO2-N", "NO3-N", "PON") & !is.na(Value)) %>%
  xtabs(~Variable + year(Time), .)
  

# Check
dat2 %>%
  filter(Variable %in% "NO3+NO2-N" & Depth1 %in% c(0,5,10) & year(Time) == 2017 & month(Time) == 2)

dat2 %>%
  mutate(Depth1 = ifelse(Depth1 == 0.5, 0, Depth1)) %>%
  filter(Variable %in% "NO3+NO2-N" & Depth1 %in% c(0, 30)) %>%
  ggplot(aes(Time, Value, color = factor(Depth1))) +
  geom_line() + geom_point() +
  facet_grid(vars(StationCode), vars(Depth1))
  

```

## 2. Nutrients, check sums
### N, plot biggest discrepancy   
- Plots all N variables (colors/circles), plus the sum of inorganic N (NH4 + NO3 + NO2) in black  
- Dotted vertical line(s) indicates where the sum is 15% higher than TOTN (example: VT67, 0 m)  
```{r}

nutrient_plot_n("VT71", 0)
nutrient_plot_n("VR51", 0)

```




### Check sums of P variables   
Check whether  
1) TotP is smaller or equal to phosphate + particular P (TOTP_P)   
2) TotP is smaller or equal to phosphate     
None are a lot off
```{r}

check_sums_p1(dat2, max_n = 1000)
check_sums_p2(dat2)

```

### P, plot biggest discrepancy   
-170%   

```{r}

nutrient_plot_p("VT71", 0)

# Doesn't have TOTP_P:
nutrient_plot_p("VR51", 0, nutrients = c("TOTP", "PO4-P"))

```



## 3. Nutrients, compare with values in Excel sheet  
  
- Compares mean values calculated from raw data with values in the "NIVAKlass" excel sheets  
- Also see the sheet 'VT71_stats' in the AqM excel file that we read in section 11  
  
### Nutrients, get mean values from raw data   
- Compare with vallues used in NIVAclass excel sheets here:  
`K:\Prosjekter\Sjøvann\KYSTOVERVÅKING ØKOKYST\KYSTOVERVÅKING ØKOKYST 2017-2020\ØKOKYST DP Norskehavet Sør I O-17091_18091_19091_200091\Rapport 2019-data\klassification`    
- VT71: Results not equal to the numbers used for VT71, but so close that none of them change class  
- VR51: Got exactly same numbers as those used
### Plot surface values for each depth/station over time  
```{r, fig.width = 8, fig.height = 5.5}

# order_as_excel <- c("VT71", "VR51")

plotdata <- dat2
nrow(plotdata)

depths <- c(0, 0.5, 5, 10)
vars <- c("TOTP", "PO4-P", "TOTN", "NO3+NO2-N", "NH4-N")
stations <- unique(plotdata$StationCode)

years <- 2017:2019
months <- list(c(6,7,8), c(1,2,12))
names(months) <- c("Summer", "Winter")

for (station in stations){
  for (i in seq_along(months)){
    df <- plotdata %>%
      mutate(Depth1 = ifelse(Depth1 == 0.5, 0, Depth1)) %>%
      mutate(Depth1 = factor(Depth1)) %>%
      filter(Variable %in% vars & 
               Depth1 %in% depths &
               month(Time) %in% months[[i]] & 
               year(Time) %in% years &
               StationCode %in% station)
    gg <- ggplot(df, aes(Time, Value, color = Depth1)) +
      geom_point() +
      geom_line() +
      facet_wrap(vars(Variable), scales = "free_y") + 
      labs(title = paste0(station, ", ", 
                         names(months)[[i]])) +
      scale_x_datetime(breaks = "1 year", date_labels = "%d.%m.%y") +
      theme(axis.text.x = element_text(hjust = 0, angle = -45))
    print(gg)
  }
}


```

### Plot weighted surface values over time   
- 4 plots, one per station and season
- Note extremely high ammonium value in VR51 winter (see section "Seasonal mean values)  
```{r, fig.width = 8, fig.height = 5.5}

# Delete two extremely high ammonium values 
nrow(dat2)
sel <- with(dat2,
            StationCode == "VR51" & 
              Variable %in% "NH4-N" &
              month(Time) == 2 &
              Value > 80)
sum(sel)

plotdata <- dat2[!sel,]
nrow(plotdata)


# order_as_excel <- c("VT71", "VR51")

months <- list(c(6,7,8), c(1,2,12))
names(months) <- c("Summer", "Winter")

for (i in seq_along(months)){
  df <- NIVAklass_get_mean_overdepth(
    years = 2017:2019,
    data = plotdata,
    vars = c("TOTP", "PO4-P", "TOTN", "NO3+NO2-N", "NH4-N"),
    months = months[[i]])
  gg <- ggplot(df, aes(Time, Value, color = StationCode)) +
    geom_point() +
    geom_line() +
    facet_wrap(vars(Variable), scales = "free_y") + 
    labs(title = names(months)[[i]]) +
    scale_x_datetime(breaks = "1 year", date_labels = "%d.%m.%y") +
    theme(axis.text.x = element_text(hjust = 0, angle = -45))
  print(gg)
}

```

### Seasonal mean values  
- Compare with vallues used in NIVAclass excel sheets here:  
`K:\Prosjekter\Sjøvann\KYSTOVERVÅKING ØKOKYST\KYSTOVERVÅKING ØKOKYST 2017-2020\ØKOKYST DP Norskehavet Sør I O-17091_18091_19091_200091\Rapport 2019-data\klassification`    
- VT71: Results not equal to the numbers used for VT71, but so close that none of them change class  
- VR51: Very samll difference between weighted and unweighted means. Got exactly same numbers as those used if we used equal weights for 0,5 and 10 meters.   
- VR51, ammonium in winter: before deleting 2 very high February values, we have unweighted mean = 49.09, in Excel 23.63. After deleting, they are quite close.  
```{r}
#
# Weights = 0.25,0.5,0.25 for 0,5 and 10 meters
#
df_mean_nutrients <- NIVAklass_nutrientvalues(2017:2019, plotdata)

df_mean_nutrients %>%
  arrange(StationCode, Season) %>%
  knitr::kable()

#
# Weights = same for 0,5 and 10 meters
#
df_mean_nutrients <- NIVAklass_nutrientvalues(2017:2019, plotdata, 
                                              depths_weights = c(1,1,1))

df_mean_nutrients %>%
  arrange(StationCode, Season) %>%
  knitr::kable()

```

## 4. Chl a quantiles    
- VR51: Fairly big difference - 3.01 used in Excel, got 3.64 here   
- VT71: Small difference - 1.78 used in Excel, got 1.72 here    
```{r}

NIVAklass_get_chl_data <- function(years, data, months = 2:10, depths = c(0,5,10)){
  data %>%
    mutate(Depth1 = ifelse(Depth1 == 0.5, 0, Depth1)) %>%
    filter(Variable %in% "KlfA" & 
             Depth1 %in% depths &
             month(Time) %in% months & year(Time) %in% years) %>% 
    select(StationCode, Time, Depth1, Variable, Value) %>%
    pivot_wider(names_from = Depth1, values_from = Value, names_prefix = "Depth")  %>%
    mutate(Max = pmax(Depth0, Depth5, Depth10, na.rm = TRUE))
    }


NIVAklass_plot_chl <- function(chldata){
  chldata %>%
    pivot_longer(c(Depth0, Depth5, Depth10, Max), names_to = "Depth", values_to = "KlfA") %>%
    ggplot(aes(Time, KlfA, color = Depth)) +
    geom_line() +
    facet_wrap(vars(StationCode))
}

NIVAklass_get_chl_quantiles <- function(chldata){
  chldata %>%
    group_by(StationCode) %>%
    summarise_at(c("Depth0", "Depth5", "Depth10", "Max"), 
             quantile, probs = 0.9, na.rm = TRUE, type = 7)  # type = 7 correponds to PERCENTILE:INC in Excel
                                                             # See part 6 above
    }


df <- NIVAklass_get_chl_data(2017:2019, dat2)

# df
NIVAklass_plot_chl(df)
NIVAklass_get_chl_quantiles(df)


```


## 5. Bottom O2    
- NOTE: Here we just check *CTD* values - the classification uses Winkler values from samples    
- Couldnæt find the lab values (at least not for 2019)  
  
Code from script 13 '13_QA_CTD_2019.Rmd'     
Remember that oxygen should be in ml/L! - See below   
```{r}
### Folders

basefolder <- "K:/Avdeling/214-Oseanografi/DATABASER/OKOKYST_2017/"

# Norskehavet Sør I
folder1 <- "OKOKYST_NH_Sor1_RMS/xlsbase/TilAquamonitor"

#
# Get data frame with one line per data file
#
# 1. Find sheet names here: (can be commented out afterwards)
# sheets_in_folder(folder1)                                 
# 2. Enter the sheet names here:  
fileinfo <- vars_in_folder(folder1, c("data","data","Data"))       

#
# Get data as list
#
datalist <- seq_len(nrow(fileinfo)) %>% map(read_data_fileno, df_fileinfo = fileinfo)

#
# Get data frame with one line per station
# NOTE: Not actally needed
#
fileinfo_stations <- seq_len(nrow(fileinfo)) %>% map_df(get_stations)
fileinfo_stations

```

### Check and combine all data   
Remember that oxygen should be in ml/L!   
```{r}

# Check variable names of each dataset
map(datalist, names)

map(datalist, ~"StationCode" %in% names(.))   # note StationId instead of StationCode in 3rd file!

# Check start of file of each dataset
map(datalist, head)  

# For file no 3, we want Oxygen__1! (oxygen in ml/L)
#   (inspect 2nd row of Excel file)
# So we delete the original Oksygen and make a new one
datalist[[3]] <- datalist[[3]] %>%
  select(-Oksygen) %>%
  rename(Oksygen = Oksygen__1) %>%
  rename(StationCode = StationId)    # Also remember this!

# datalist[[3]] %>% filter(Depth1 == 200.5) %>% View("Data no 3")

# Combine
data_all <- bind_rows(datalist)

data_all <- data_all %>%
  filter(Oksygen > 0)

# Check
# data_all %>% filter(Depth1 == 200.5) %>% View("Data_all")

```


#### Data availability and max depth  
```{r}

for (station in c("VR51", "VT71")){
  
  dat <- data_all %>%
    filter(StationCode %in% station)
  
  # Plot data availablity 1 - number of data with/without Oxygen
  gg <- dat %>%
    count(Date, Has_data = !is.na(Oksygen)) %>%
    ggplot(aes(Date, n, color = Has_data)) + 
    geom_vline(aes(xintercept = Date), color = "grey60") +
    geom_point() +
    labs(title = paste(station, "- number of data with/without oxygen"))
  print(gg)
  
  # Plot data availablity 2 - max depth
  gg <- dat %>%
    filter(!is.na(Oksygen)) %>%
    group_by(Date) %>%
    summarise(Max_depth = max(Depth1)) %>%
    ggplot(aes(Date, Max_depth)) + 
    geom_vline(aes(xintercept = Date), color = "grey60")+
    geom_point() +
    labs(title = paste(station, "- max depth of oxygen data"))
  print(gg)
  
}

```

#### Get oxygen at max depth    
Also calculates 'Max_depth_perc' = Max depth that date as percent of max depth of that station
```{r}
# Get data from largest depth
# 'Max_depth_perc'
dat_bottom <- data_all %>%
  filter(!is.na(Oksygen)) %>%
  group_by(StationCode) %>%
  mutate(Max_depth_station = max(Depth1)) %>%
  group_by(StationCode, Date) %>%
  mutate(Max_depth = max(Depth1),
         Max_depth_perc = Max_depth/Max_depth_station*100) %>%
  filter(Depth1 == Max_depth) # 230 is from plot above

# Check 
dat_bottom %>%
  select(StationCode, Date, Oksygen, Max_depth, Max_depth_station, Max_depth_perc)

# Keep only those where max depth is 85% of depth at station
dat_bottom <- dat_bottom %>%
  filter(Max_depth_perc > 85)

```

#### Get oxygen at oxygen minimum
```{r}
dat_o2min <- data_all %>%
  filter(!is.na(Oksygen)) %>%
  group_by(StationCode, Date) %>%
  mutate(Min_O2 = min(Oksygen)) %>%
  filter(Oksygen == Min_O2) %>% # select(Date, Oksygen, Depth1)
  summarise(Oksygen = mean(Oksygen), 
            Oksygenmetning = mean(Oksygenmetning), 
            Depth1_min = min(Depth1),
            Depth1_max = max(Depth1),
            Depth1 = max(Depth1)
            ) %>%
  mutate(Depth_diff = Depth1_max - Depth1_min)

# Check those where mn oxygen occurs over > 2 meter depth
dat_o2min %>% 
  filter(Depth_diff > 2) %>%
  select(StationCode, Date, Oksygen, Depth1_min, Depth1_max)

```

#### Combine oxygen data  
```{r}
dat_oxygen <- bind_rows(
  dat_bottom %>% 
    select(StationCode, Date, Oksygen, Oksygenmetning, Depth1) %>% 
    mutate(Position = "Bottom"),
  dat_o2min %>% 
    select(StationCode, Date, Oksygen, Oksygenmetning, Depth1) %>% 
    mutate(Position = "O2-minimum")
  ) %>%
  ungroup()

dat_oxygen

```

#### Plot time series   
Shading for winter season (October - April)
```{r}

for (station in c("VR51", "VT71")){
  gg1 <- plot_ts(
    data = dat_oxygen %>% filter(StationCode %in% station), 
    years = 2017:2019, 
    y_variable = "Oksygen", category_variable = "Position", 
    quality_type = "Concentration",
    title = paste(station, "- oxygen concentration"))
  gg2 <- plot_ts(
    data = dat_oxygen %>% filter(StationCode %in% station), 
    years = 2017:2019, 
    category_variable = "Position",
    y_variable = "Oksygenmetning", 
    quality_type = "Saturation",
    title = paste(station, "- oxygen saturation"))
  print(gg1)
  print(gg2)
}


```

#### Make means  
Mean for last 3 years = 3.844558    
Mean for last 3 years, winer only = 3.641219   

```{r, results = 'asis'}
# Mean oxygen for all months
dat_oxygen %>%
  group_by(StationCode, Position) %>%
  summarise(Oksygen = mean(Oksygen), 
            Oksygenmetning = mean(Oksygenmetning)) %>%
  knitr::kable(caption = "Mean oxygen for all months", digits = 2)

# Mean oxygen for winter (Sept - April)
dat_oxygen %>%
  filter(month(Date) <= 4 | month(Date) >= 9) %>%
  group_by(StationCode, Position) %>%
  summarise(Oksygen = mean(Oksygen),
            Oksygenmetning = mean(Oksygenmetning)) %>%
  knitr::kable(caption = "Mean oxygen for winter (Sept - April)", digits = 2)

```

