---
title: "QC Norskehavet Sør I, nutrients + Chl"
author: "DHJ"
date: "28.02.2021"
output: 
  html_document:
    toc: true
    toc_float: true
    keep_md: true
---

**QC of Norskehavet Sør I (Caroline + Lars)**      
- Checked input to these files:  
`K:\Prosjekter\Sjøvann\KYSTOVERVÅKING ØKOKYST\KYSTOVERVÅKING ØKOKYST 2017-2020\ØKOKYST DP Norskehavet Sør I O-17091_18091_19091_200091\Rapport 2019-data\klassification`
- Stations:
    + VT71 Skinnbrokleia
    + VR51 Korsen
  
- **Note: Here we download data from Aquamonitor insted of reading excel files (as last year)**  

**Veileder:**  
- `Documents\artikler\veiledere etc\02_2013_klassifiserings-veileder_.pdf`  
- Chlorophyll: page 91.   
"I SørNorge (til Stadt) anbefales det at innsamlingen starter i februar og avsluttes ved utgangen av oktober."   
"Nord for Stadt anbefales det at innsamlingsperioden strekker seg fra mars til og med september."   
90% percentiles should be used   
  
- Also see calculation example on on page 187-189, and the comment on page 189:  
"Det gjøres oppmerksom på at dataprogrammene som beregner 90-persentil, gjerne benytter ulike metoder". We use the Excel method PERCENTILE.EXC, corresponding to type 6 quantiles in R  
  
- Nutrients: page 102 in veileder  
Vinter er november til og med februar   
Sommer er juni til og med august  

  
```{r, echo = FALSE, results='hide', warning=FALSE, message=FALSE}

### 0. Libraries   

library(dplyr)
library(purrr)
library(ggplot2)
library(lubridate)
library(readxl)
library(tidyr)
library(knitr)         
library(RColorBrewer)

# library(niRvana)

source('11_Read_from_Nivabasen_functions.R')  # Used to read from Nivadatabase

source("12_QA_2019_from_excel_functions.R")
source("13_QA_CTD_2019_functions.R")          # plot_ctdprofile_station

library(niRvana)

knitr::opts_chunk$set(results = 'hold', echo = FALSE)

# RColorBrewer::display.brewer.all()
```


```{r, echo = FALSE}

### Set Nivadabase password 

# Do this once:
# set_credentials()

```


## 1. Read Norskehavet Nord I nutrients     

```{r, echo = FALSE}

### Test  
# - Run only to find which parameter names/ids to use (from WC_PARAMETER_DEFINITIONS)   

# We start with station VR51, and dig all the way from 
# table PROJECTS_STATIONS to WC_PARAMETER_DEFINITIONS

if (FALSE){
  
  # For copy/pasting column names  
  nm <- function(df)
    names(df)  %>% paste(collapse = ", ")
  
  df_projstat <- get_nivabase_selection(
    "PROJECT_ID, STATION_ID, STATION_CODE, STATION_NAME, STATION_IS_ACTIVE", 
    "PROJECTS_STATIONS", 
    "STATION_CODE", 
    "VR51", values_are_text = TRUE)
  df_projstat
  
  df_watersamp <- get_nivabase_selection(
    "WATER_SAMPLE_ID, STATION_ID, SAMPLE_DATE, DEPTH1, DEPTH2, REMARK, SAMPLE_POINT_ID",
    "WATER_SAMPLES", 
    "STATION_ID, ", 
    df_projstat$STATION_ID)
  
  df_waterchem <- get_nivabase_selection(
    "WATER_SAMPLE_ID, METHOD_ID, VALUE, UNCERTAINTY, FLAG1, FLAG2, REMARK, APPROVED", 
    "WATER_CHEMISTRY_VALUES", 
    "WATER_SAMPLE_ID", 
    subset(df_watersamp, year(SAMPLE_DATE) == 2020 & month(SAMPLE_DATE) == 8)$WATER_SAMPLE_ID)
  
  # df_waterchem %>% nm() 
  
  xtabs(~METHOD_ID, df3a)
  
  df_m <- get_nivabase_selection(
    "METHOD_ID, NAME, UNIT, LABORATORY, DESCR, MATRIX, CAS, IUPAC", 
    "METHOD_DEFINITIONS", 
    "METHOD_ID", 
    unique(df_waterchem$METHOD_ID))
  
  df_mpar  <- get_nivabase_selection(
    "PARAMETER_ID, METHOD_ID, CONVERSION_FACTOR", 
    "WC_PARAMETERS_METHODS", 
    "METHOD_ID", 
    df_m$METHOD_ID)
  # nm(df_mpar)
  
  df_par <- get_nivabase_selection(
    "*", 
    "WC_PARAMETER_DEFINITIONS", 
    "PARAMETER_ID", 
    unique(df_mpar$PARAMETER_ID))
  par_names1 <- df_par$NAME
  
  # Get the other TOTN parameters too
  df_par_all <- get_nivabase_data(
    "select * from NIVADATABASE.WC_PARAMETER_DEFINITIONS")  # 716 rows
  nm(df_par_all)
  
  par_names2 <- df_par_all %>%
    filter(grepl("TOTN", NAME) & !DESCR %in% "FAGDATA2") %>%
    pull(NAME)
  
  par_names <- unique(c(par_names1, par_names2))
  par_names
  
} 
# dput(par_names)

```

### Get data  
```{r}

# Read from NIvabase (1 minute) or read from saved data?
read_from_nivabase <- FALSE


par_names <- c("NO3+NO2-N", "NH4-N", "TOTP", "PO4-P", "SiO2", 
               "KlfA", "TSM", 
               "TOTN (old EF)", "TOTN", "TOTN (est.)")

stations <-  c("VR51", "VT71")


if (read_from_nivabase){
  # debugonce(get_water_chemistry)
  df_chem <- get_water_chemistry(stationcodes = stations,
                                 parameters = par_names, 
                                 years = 2017:2020) %>%
    mutate(Year = year(SAMPLE_DATE),
           Month = month(SAMPLE_DATE))
  
  # Uncomment to save again:
  # saveRDS(df_chem, "Data/17_2020_df_chem.rds")  
  
} else {

  df_chem <- readRDS("Data/17_2020_df_chem.rds")  
  
}

cat(nrow(df_chem), "rows of data downloaded \n")
cat("Stations:", paste(stations, collapse = ", "), "\n")
cat("Parameters:", paste(par_names, collapse = ", "), "\n")

```


### Measurements by parameter, site and year    
```{r}

xtabs(~Year + PARAM_NAME + STATION_CODE, df_chem)

```
### Measurements by depth, site and year    
```{r}

df_chem %>%
  distinct(STATION_CODE, Year, DEPTH1, SAMPLE_DATE) %>%
  xtabs(~ DEPTH1 + Year + STATION_CODE, .)

```
### Parameters  
```{r}

check <- df_chem %>%
  group_by(STATION_CODE, SAMPLE_DATE, DEPTH1, PARAM_NAME) %>%
  summarise(n = n(), .groups = 'drop') %>%
  filter(n > 1)

if (nrow(check) > 0)
  stop("Some combinations of STATION_CODE, SAMPLE_DATE, DEPTH1, PARAM_NAME occurs >1 time!")

df_chem %>%
  count(PARAM_NAME, NAME)

```

### Reformat data    
Formatting the data in 'wide' format (one column per paremeter) as shown below  
```{r}

dat <- df_chem %>%
  select(STATION_CODE, SAMPLE_DATE, DEPTH1, PARAM_NAME, VALUE) %>%
  tidyr::pivot_wider(names_from = "PARAM_NAME", values_from = "VALUE") %>%
  rename(StationCode = STATION_CODE, 
         Date = SAMPLE_DATE, 
         Depth1 = DEPTH1) %>%
  mutate(Year = year(Date),
         Month = month(Date),  
         `Nitrate + ammonium` = `NO3+NO2-N` + `NH4-N`)  


head(dat, 4)

```
  
## Chl a   
```{r, echo = FALSE, fig.width=10, fig.height=4.5, warning=FALSE}

stations %>% walk(plot_ctdprofile_station, data = dat, 
                  variable = "KlfA", points = TRUE, year_by_month = TRUE)

```

## Nutrients  {.tabset}

```{r, echo = FALSE, fig.width=10, fig.height=4.5, warning=FALSE}

# 
# Alternative way, prints all variables + stations - but doesn't give headings  
#
if (FALSE){
  
  df <- df_chem %>%
    distinct(STATION_CODE, PARAM_NAME) 
  
  walk2(
    df$STATION_CODE[1:3],
    df$PARAM_NAME[1:3],
    ~plot_ctdprofile_station(stationcode = .x, 
                             variable = .y,
                             data = dat, 
                             points = TRUE, year_by_month = TRUE)
  )
  
}

```

### NH4-N
```{r, echo = FALSE, fig.width=10, fig.height=4.5, warning=FALSE}

stations %>% walk(plot_ctdprofile_station, data = dat, 
                  variable = "NH4-N", points = TRUE, year_by_month = TRUE)

```


### NO3+NO2-N
```{r, echo = FALSE, fig.width=10, fig.height=4.5, warning=FALSE}

stations %>% walk(plot_ctdprofile_station, data = dat, 
                  variable = "NO3+NO2-N", points = TRUE, year_by_month = TRUE)

```


### TOTN
```{r, echo = FALSE, fig.width=10, fig.height=4.5, warning=FALSE}

stations %>% walk(plot_ctdprofile_station, data = dat, 
                  variable = "TOTN", points = TRUE, year_by_month = TRUE)

```


### TOTN (old EF) 
```{r, echo = FALSE, fig.width=10, fig.height=4.5, warning=FALSE}

stations %>% walk(plot_ctdprofile_station, data = dat, 
                  variable = "TOTN (old EF)", points = TRUE, year_by_month = TRUE)

```


### PO4-P
```{r, echo = FALSE, fig.width=10, fig.height=4.5, warning=FALSE}

stations %>% walk(plot_ctdprofile_station, data = dat, 
                  variable = "PO4-P", points = TRUE, year_by_month = TRUE)

```

### TOTP
```{r, echo = FALSE, fig.width=10, fig.height=4.5, warning=FALSE}

stations %>% walk(plot_ctdprofile_station, data = dat, 
                  variable = "TOTP", points = TRUE, year_by_month = TRUE)

```

### SiO2
```{r, echo = FALSE, fig.width=10, fig.height=4.5, warning=FALSE}

stations %>% walk(plot_ctdprofile_station, data = dat, 
                  variable = "SiO2", points = TRUE, year_by_month = TRUE)

```

### TSM
```{r, echo = FALSE, fig.width=10, fig.height=4.5, warning=FALSE}

stations %>% walk(plot_ctdprofile_station, data = dat, 
                  variable = "TSM", points = TRUE, year_by_month = TRUE)

```

## TotN versus the sum of nitrate + ammonium    
```{r, echo = FALSE, fig.width=10, fig.height=5.5, warning=FALSE}

df_nit <- dat %>%
  select(StationCode, Year, Month, Depth1, `Nitrate + ammonium`, TOTN, `TOTN (old EF)`) %>%
  tidyr::pivot_longer(
    `Nitrate + ammonium`:`TOTN (old EF)`,
    names_to = "Parameter",
    values_to = "Concentration")

func <- function(station){
  
  df_nit %>%
    filter(StationCode == station) %>%
    ggplot(aes(Concentration, Depth1, color = Parameter)) +
    geom_path() + geom_point() +
    scale_y_reverse() +
    facet_grid(rows = vars(Year), cols = vars(Month)) +
    theme(legend.position = "bottom",
          axis.text.x = element_text(angle = -45, hjust = 0)) +
    labs(title = paste0(station, ", TotN versus the sum of nitrate + ammonium"))

}

func(stations[[1]])
func(stations[[2]])

```




