---
title: "QC Norskehavet Sør 2, nutrients + Chl"
author: "DHJ"
date: "28.02.2021"
output: 
  html_document:
    toc: true
    toc_float: true
    keep_md: true
---

**QC of Norskehavet Sør I (Caroline + Lars)**      
- Stations:  
    + VR52: OKOKYST_NH_Sor2_Aquakompetanse  
    + VT42: OKOKYST_NH_Sor2_MagneAuren   
    + VR31: OKOKYST_NH_Sor2_SNO:   
    
- **Note: Here we download data from Aquamonitor insted of reading excel files (as last year)**  

**Veileder:**  
- `Documents\artikler\veiledere etc\02_2013_klassifiserings-veileder_.pdf`  
- Chlorophyll: page 91.   
"I SørNorge (til Stadt) anbefales det at innsamlingen starter i februar og avsluttes ved utgangen av oktober."   
"Nord for Stadt anbefales det at innsamlingsperioden strekker seg fra mars til og med september."   
90% percentiles should be used   
  
- Also see calculation example on on page 187-189, and the comment on page 189:  
"Det gjøres oppmerksom på at dataprogrammene som beregner 90-persentil, gjerne benytter ulike metoder". We use the Excel method PERCENTILE.EXC, corresponding to type 6 quantiles in R  
  
- Nutrients: page 102 in veileder  
Vinter er november til og med februar   
Sommer er juni til og med august  

  
```{r, echo = FALSE, results='hide', warning=FALSE, message=FALSE}

### 0. Libraries   

library(dplyr)
library(purrr)
library(ggplot2)
library(lubridate)
library(readxl)
library(tidyr)
library(knitr)         
library(RColorBrewer)

# library(niRvana)

source('11_Read_from_Nivabasen_functions.R')  # Used to read from Nivadatabase

source("12_QA_2019_from_excel_functions.R")
source("13_QA_CTD_2019_functions.R")          # plot_ctdprofile_station

library(niRvana)

knitr::opts_chunk$set(results = 'hold', echo = FALSE)

# RColorBrewer::display.brewer.all()
```


```{r, echo = FALSE}

### Set Nivadabase password 

# Do this once:
# set_credentials()

```


## 1. Read Norskehavet Nord I nutrients     

### Get data  
```{r}

# Read from NIvabase (1 minute) or read from saved data?
read_from_nivabase <- FALSE
# read_from_nivabase <- TRUE


par_names <- c("NO3+NO2-N", "NH4-N", "TOTP", "PO4-P", "SiO2", 
               "KlfA", "TSM", 
               "TOTN (old EF)", "TOTN", "TOTN (est.)")

stations <-  c("VR52", "VT42", "VR31")


if (read_from_nivabase){
  # debugonce(get_water_chemistry)
  df_chem <- get_water_chemistry(stationcodes = stations,
                                 parameters = par_names, 
                                 years = 2017:2020) %>%
    mutate(Year = year(SAMPLE_DATE),
           Month = month(SAMPLE_DATE))
  
  # Uncomment to save (overwrite) data:
  # saveRDS(df_chem, "Data/17_2020_NorskehavetS2.rds")  
  
} else {

  df_chem <- readRDS("Data/17_2020_NorskehavetS2.rds")  
  
}

cat(nrow(df_chem), "rows of data downloaded \n")
cat("Stations:", paste(stations, collapse = ", "), "\n")
cat("Parameters:", paste(par_names, collapse = ", "), "\n")

```


### Measurements by parameter, site and year    
```{r}

xtabs(~Year + PARAM_NAME + STATION_CODE, df_chem)

```
### Measurements by depth, site and year    
```{r}

df_chem %>%
  distinct(STATION_CODE, Year, DEPTH1, SAMPLE_DATE) %>%
  xtabs(~ DEPTH1 + Year + STATION_CODE, .)

```
### Parameters  
```{r}

check <- df_chem %>%
  group_by(STATION_CODE, SAMPLE_DATE, DEPTH1, PARAM_NAME) %>%
  summarise(n = n(), .groups = 'drop') %>%
  filter(n > 1)

if (nrow(check) > 0)
  stop("Some combinations of STATION_CODE, SAMPLE_DATE, DEPTH1, PARAM_NAME occurs >1 time!")

df_chem %>%
  count(PARAM_NAME, NAME)

```

### Reformat data    
Formatting the data in 'wide' format (one column per paremeter) as shown below  
```{r}

dat <- df_chem %>%
  select(STATION_CODE, SAMPLE_DATE, DEPTH1, PARAM_NAME, VALUE) %>%
  tidyr::pivot_wider(names_from = "PARAM_NAME", values_from = "VALUE") %>%
  rename(StationCode = STATION_CODE, 
         Date = SAMPLE_DATE, 
         Depth1 = DEPTH1) %>%
  mutate(Year = year(Date),
         Month = month(Date),  
         `Nitrate + ammonium` = `NO3+NO2-N` + `NH4-N`) %>%
  arrange(StationCode, Date, Depth1)


head(dat, 4)

```
  
## Chl a   
```{r, echo = FALSE, fig.width=10, fig.height=4.5, warning=FALSE}

stations %>% walk(plot_ctdprofile_station, data = dat, 
                  variable = "KlfA", points = TRUE, year_by_month = TRUE)

```

## Nutrients  {.tabset}

### NH4-N
```{r, echo = FALSE, fig.width=10, fig.height=4.5, warning=FALSE}

stations %>% walk(plot_ctdprofile_station, data = dat, 
                  variable = "NH4-N", points = TRUE, year_by_month = TRUE)

```


### NO3+NO2-N
```{r, echo = FALSE, fig.width=10, fig.height=4.5, warning=FALSE}

stations %>% walk(plot_ctdprofile_station, data = dat, 
                  variable = "NO3+NO2-N", points = TRUE, year_by_month = TRUE)

```


### TOTN
```{r, echo = FALSE, fig.width=10, fig.height=4.5, warning=FALSE}

stations %>% walk(plot_ctdprofile_station, data = dat, 
                  variable = "TOTN", points = TRUE, year_by_month = TRUE)

```


### TOTN (old EF) 
```{r, echo = FALSE, fig.width=10, fig.height=4.5, warning=FALSE}

stations %>% walk(plot_ctdprofile_station, data = dat, 
                  variable = "TOTN (old EF)", points = TRUE, year_by_month = TRUE)

```


### PO4-P
```{r, echo = FALSE, fig.width=10, fig.height=4.5, warning=FALSE}

stations %>% walk(plot_ctdprofile_station, data = dat, 
                  variable = "PO4-P", points = TRUE, year_by_month = TRUE)

```

### TOTP
```{r, echo = FALSE, fig.width=10, fig.height=4.5, warning=FALSE}

stations %>% walk(plot_ctdprofile_station, data = dat, 
                  variable = "TOTP", points = TRUE, year_by_month = TRUE)

```

### SiO2
```{r, echo = FALSE, fig.width=10, fig.height=4.5, warning=FALSE}

stations %>% walk(plot_ctdprofile_station, data = dat, 
                  variable = "SiO2", points = TRUE, year_by_month = TRUE)

```

### TSM
```{r, echo = FALSE, fig.width=10, fig.height=4.5, warning=FALSE}

stations %>% walk(plot_ctdprofile_station, data = dat, 
                  variable = "TSM", points = TRUE, year_by_month = TRUE)

```

## TotN versus the sum of nitrate + ammonium    
```{r, echo = FALSE, fig.width=10, fig.height=5.5, warning=FALSE}

df_nit <- dat %>%
  select(StationCode, Year, Month, Depth1, `Nitrate + ammonium`, TOTN, `TOTN (old EF)`) %>%
  tidyr::pivot_longer(
    `Nitrate + ammonium`:`TOTN (old EF)`,
    names_to = "Parameter",
    values_to = "Concentration")

func <- function(station){
  
  gg <- df_nit %>%
    filter(StationCode == station) %>%
    ggplot(aes(Concentration, Depth1, color = Parameter)) +
    geom_path() + geom_point() +
    scale_y_reverse() +
    facet_grid(rows = vars(Year), cols = vars(Month)) +
    theme(legend.position = "bottom",
          axis.text.x = element_text(angle = -45, hjust = 0)) +
    labs(title = paste0(station, ", TotN versus the sum of nitrate + ammonium"))
  print(gg)
  
}

stations %>% walk(func)

```




