---
title: "Match ferrybox sample numbers to stations"
author: "DHJ"
date: "17 februar 2020"
output: 
  html_document:
    toc: true    
    toc_flo: true
  github_document:
    html_preview: false
    toc: true
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_dir = "output", output_format = "all") })    

---

Find which sample numbers (in Ferrybox files) belongs to which stations  
    
Corresponding numbers in Fantasy:    
- VT4 = sample no 21-22  
    
Often, but not always (!), corresponding numbers in Trollfjord:    
- VT4 = sample no 23-24    
- VT72 = sample no 21-22     
- VT23 = sample no 17-18    
- VT80 = sample no 11-12  
- VT45 = sample no 13-14  
- VT22 = sample no 15-16   
- VR25 = sample no 6-7   
- VR23 = sample no 4-5  
- VR76 = sample no 2-3  
For Trollfjord, we make a search for closest station as well (part 5 below)
  
Ferrybox sample files for 2018-2019 saved as  
- "Datasett/21_df_ferrybox_sampledata_fa.rds"    
- "Datasett/21_df_ferrybox_sampledata_tf.rds"  
- "Datasett/Ferrybox_samples_OneDrive/FA_2018-19_automatic_samples.xlsx"   
- "Datasett/Ferrybox_samples_OneDrive/TF_2018-19_automatic_samples.xlsx"   


## 1. Libraries
```{r}

library(plyr)      # used by functions in "Get_files_NIVA_ftp_server_functions.R" 
library(dplyr)     # load AFTER plyr, so e,g, we use count() from dplyr
library(purrr)
library(ggplot2)
library(lubridate)
library(readxl)
library(tidyr)
library(RColorBrewer)
# For mapping:
library(mapview)
library(sf)

source("Get_files_NIVA_ftp_server_functions.R")

# library(niRvana)
#source("12_QA_2019_from_excel_functions.R")
# RColorBrewer::display.brewer.all()
```

## 2. Read AqM station data
```{r}
df_aqm_stations <- read_excel("Datasett/AqM_2017_2019_ØKOKYST_Ferrybox_ToR.xlsx", 
                              sheet = "StationPoint")

```


## 3. Ferrybox log data 2019

### a. Ferrybox log data 2019
We do this before 2018 in order to get the headers (column names),
which we will use for 2018 data    
- Data downloaded from Pierre's OneDrive 12.02.2020    
- [Link to Pierre's OneDrive](https://niva365-my.sharepoint.com/:f:/g/personal/pierre_jaccard_niva_no/Et9g2CZL2_tEg3vOt7ACQh0BHCjrOcKzPz-iAgdYst2ddQ?e=hpkNeW)  

```{r}

df_ferrybox_sampledata_2019_fa <- readxl::read_excel(
  "Datasett/Ferrybox_samples_OneDrive/FA_2019_automatic_samples.xlsx")

df_ferrybox_sampledata_2019_tf <- readxl::read_excel(
  "Datasett/Ferrybox_samples_OneDrive/TF_2019_automatic_samples.xlsx")

# names(df_ferrybox_sampledata_2019_fa)
# names(df_ferrybox_sampledata_2019_tf)

X <- df_ferrybox_sampledata_2019_fa$SYSTEM_DATE_DMY - ymd_hms("1899-12-31 00:00:00")

df_ferrybox_sampledata_2019_fa <- df_ferrybox_sampledata_2019_fa %>%
  mutate(TIME = SYSTEM_TIME + X)

# Check
# df_samples_fa$SYSTEM_DATE_DMY %>% head()
# df_samples_fa$SYSTEM_TIME %>% head()
# df_samples_fa$TIME %>% head()

unique(df_ferrybox_sampledata_2019_fa$SYSTEM_DATE_DMY)

```

### b. Ferrybox log data 2018   
```{r}

# 4. Download all Ferrybox log data    
# Or just use saved log data

redownload_data <- FALSE  # if FALSE, we just read the saved data


if (redownload_data){  # takes just a couple of minutes
  
  #
  # a. Get dates and filenames ('df_samplefilenames_fa_2018')  
  #
  # get_folders()
  
  #
  # Fantasy
  #
  
  # Get the url for the vessel we want
  url.folder <- get_url_foldername("fantasy")
  
  # Get file names of all log files for that vessel
  samplefiles_fn <- get_filenames_samplefiles("fantasy")
  
  head(samplefiles_fn)
  # Check
  # tail(logfiles_fn)
  
  # Combine with dates 
  df_samplefilenames_fa_2018 <- data.frame(
    filename = samplefiles_fn,
    Date = ymd(substr(samplefiles_fn, 9, 16)),   # note that 9,16 is specifically for 'samples'
    stringsAsFactors = FALSE) %>%
    filter(year(Date) == 2018)
  
  
  #
  # Trollfjord
  #
  
  # Get the url for the vessel we want
  url.folder <- get_url_foldername("trollfjord")
  
  # Get file names of all log files for that vessel
  samplefiles_fn <- get_filenames_samplefiles("trollfjord")
  
  # Check
  # tail(logfiles_fn)
  
  # Combine with dates 
  df_samplefilenames_tf_2018 <- data.frame(
    filename = samplefiles_fn,
    Date = ymd(substr(samplefiles_fn, 9, 16)), # note that 9,16 is specifically for 'samples'
    stringsAsFactors = FALSE) %>%
    filter(year(Date) == 2018)
  
  #
  # b. Combine filenames  
  #
  
  df_samplefilenames_2018 <- 
    bind_rows(
      df_samplefilenames_fa_2018 %>% mutate(Ferrybox = "fantasy"),
      df_samplefilenames_tf_2018 %>% mutate(Ferrybox = "trollfjord")
    )
  
  #
  # c. Download sample file data
  #
  df_ferrybox_sampledata_2018_fa <- get_several_logfiles(
    df_samplefilenames_fa_2018$filename, 
    vessel = "fantasy",
    trace = FALSE)
  
  df_ferrybox_sampledata_2018_tf <- get_several_logfiles(
    df_samplefilenames_tf_2018$filename, 
    vessel = "trollfjord",
    trace = FALSE)
  
  names(df_ferrybox_sampledata_2018_fa) <- names(df_ferrybox_sampledata_2019_fa)
  names(df_ferrybox_sampledata_2018_tf) <- names(df_ferrybox_sampledata_2019_tf)

  #
  # d. Save
  #
  saveRDS(df_ferrybox_sampledata_2018_fa,
          "Datasett/21_df_ferrybox_sampledata_2018_fa.rds")
  saveRDS(df_ferrybox_sampledata_2018_tf,
          "Datasett/21_df_ferrybox_sampledata_2018_tf.rds")
  
  
} else {
  
  df_ferrybox_sampledata_2018_fa <-
    readRDS("Datasett/21_df_ferrybox_sampledata_2018_fa.rds")
  df_ferrybox_sampledata_2018_tf <-
    readRDS("Datasett/21_df_ferrybox_sampledata_2018_tf.rds")
  
}


```


### c. Combine data for both years for each ferry     
- 2018 + 2019  
- The file columns differ betrween ferries, so we donæt combine them  
```{r}

#
# Fantasy
#

# df_ferrybox_sampledata_2018_fa[1:7] %>% str()
# df_ferrybox_sampledata_2019_fa[1:7] %>% str()

df_ferrybox_sampledata_2018_fa <- df_ferrybox_sampledata_2018_fa %>%
  mutate(SYSTEM_DATE_DMY = dmy_hms(paste(SYSTEM_DATE_DMY, "00:00:00")),
         SYSTEM_TIME = ymd_hms(paste("1899-12-31", SYSTEM_TIME)))

df_ferrybox_sampledata_fa <- bind_rows(
  df_ferrybox_sampledata_2018_fa,
  df_ferrybox_sampledata_2019_fa)

#
# Trollfjord
#
# df_ferrybox_sampledata_2018_tf[1:7] %>% str()
# df_ferrybox_sampledata_2019_tf[1:7] %>% str()

df_ferrybox_sampledata_2018_tf <- df_ferrybox_sampledata_2018_tf %>%
  mutate(SYSTEM_DATE = dmy_hms(paste(SYSTEM_DATE, "00:00:00")),
         SYSTEM_TIME = ymd_hms(paste("1899-12-31", SYSTEM_TIME)))

df_ferrybox_sampledata_tf <- bind_rows(
  df_ferrybox_sampledata_2018_tf,
  df_ferrybox_sampledata_2019_tf)

# names(df_ferrybox_sampledata_fa)
# names(df_ferrybox_sampledata_tf)

# df_ferrybox_sampledata <- 
#   bind_rows()



```

### d. Save sample files  
```{r}

saveRDS(df_ferrybox_sampledata_fa,
        "Datasett/21_df_ferrybox_sampledata_fa.rds")
saveRDS(df_ferrybox_sampledata_tf,
        "Datasett/21_df_ferrybox_sampledata_tf.rds")

openxlsx::write.xlsx(df_ferrybox_sampledata_fa,
                     "Datasett/Ferrybox_samples_OneDrive/FA_2018-19_automatic_samples.xlsx")
openxlsx::write.xlsx(df_ferrybox_sampledata_tf,
                     "Datasett/Ferrybox_samples_OneDrive/TF_2018-19_automatic_samples.xlsx")
  
```



### e. Sample numbers used in files   
Fantasy:     
- Mostly MANUAL_SAMPLE_COUNTER = 0 and AUTOMATIC_SAMPLE_COUNTER = 1-24   
- For a few lines MANUAL_SAMPLE_COUNTER = 1-24 and AUTOMATIC_SAMPLE_COUNTER = 0   
Trollfjord:     
- Same but variables are called SAMPLE_MAN_NUM and SAMPLE_AUTO_NUM   
```{r, results = 'hold'}

cat("================================================\n")
cat("     FANTASY\n")
#
# Fantasy
#
cat("----------------------------\n")
cat("MANUAL_SAMPLE_COUNTER == 0\n")
df_ferrybox_sampledata_fa %>% 
  filter(MANUAL_SAMPLE_COUNTER == 0) %>% 
  xtabs(~AUTOMATIC_SAMPLE_COUNTER, .)

cat("----------------------------\n")
cat("AUTOMATIC_SAMPLE_COUNTER == 0\n")
df_ferrybox_sampledata_fa %>% 
  filter(AUTOMATIC_SAMPLE_COUNTER == 0) %>% 
  xtabs(~MANUAL_SAMPLE_COUNTER, .)

cat("----------------------------\n")
cat("Both == 0\n")
df_ferrybox_sampledata_fa %>% 
  filter(MANUAL_SAMPLE_COUNTER == 0 & AUTOMATIC_SAMPLE_COUNTER == 0) %>% nrow()

cat("================================================\n")
cat("     TROLLFJORD\n")
#
# Trollfjord
#

cat("----------------------------\n")
cat("SAMPLE_MAN_NUM == 0\n")
df_ferrybox_sampledata_tf %>% 
  filter(SAMPLE_MAN_NUM == 0) %>%
  xtabs(~SAMPLE_AUTO_NUM, .)

cat("----------------------------\n")
cat("SAMPLE_AUTO_NUM == 0\n")
df_ferrybox_sampledata_tf %>% 
  filter(SAMPLE_AUTO_NUM == 0) %>%
  xtabs(~SAMPLE_MAN_NUM, .)

```

## 4. Match automatic samples    
Find which sample numbers (in Ferrybox files) belongs to which stations  

### Set time and direction for Trollfjord
```{r}

df_ferrybox_sampledata_tf <- df_ferrybox_sampledata_tf %>%
  mutate(
    x1 = as.character(SYSTEM_DATE),
    x2 = as.character(SYSTEM_TIME) %>% substr(12,19),
    Date = ymd(x1),
    Time = ymd_hms(paste(x1, x2))) %>%   # select(x1, x2, Date, Time)
  select(-x1, -x2, -SYSTEM_TIME, -SYSTEM_DATE) %>%
  arrange(Time) %>%
  mutate(
    dhour = (as.numeric(Time) - lag(as.numeric(Time), order_by = Time))/3600,
    dlat = GPS_LAT - lag(GPS_LAT, order_by = Time),
    dlon = GPS_LON - lag(GPS_LON, order_by = Time),
    ) %>%  #  elect(Date, Time, GPS_LAT, dhour, dlat)
  select(SHIP_CODE, Date, Time, dhour, dlat, dlon, SAMPLE_AUTO_NUM, SAMPLE_MAN_NUM, TRIP, everything())

```



### Prepare ferrybox samples data for mapping    
```{r}


#
# Fantasy (sf_points_fa)
#

# Set rownames to "FullStationName" - these will show up when you hover over points with the mouse
df_points <- df_ferrybox_sampledata_fa %>%
  filter(AUTOMATIC_SAMPLE_COUNTER > 0) %>%
  # "copy" coordinates so the show up in popup
  mutate(Lon = GPS_LONGITUDE, Lat = GPS_LATITUDE) %>%  
  as.data.frame()   # read_excel makes a tibble, but you can't set row names for tibbles 

# Set rownames (shown at mouse hover) - must be unique
rownames(df_points) <- with(df_points,
                            paste(SHIP_CODE, AUTOMATIC_SAMPLE_COUNTER, 
                                  SYSTEM_DATE_DMY))

sf_points_fa <- st_as_sf(df_points,
                          coords = c("GPS_LONGITUDE", "GPS_LATITUDE"),
                          crs = "+proj=longlat +ellps=WGS84")

#
# Trollfjord (sf_points_tf)
#

# Set rownames to "FullStationName" - these will show up when you hover over points with the mouse
df_points <- df_ferrybox_sampledata_tf %>%
  filter(SAMPLE_AUTO_NUM > 0) %>%
  # "copy" coordinates so the show up in popup
  mutate(Lon = GPS_LON, Lat = GPS_LAT) %>%  
  as.data.frame()   # read_excel makes a tibble, but you can't set row names for tibbles 

# Set rownames (shown at mouse hover) - must be unique
rownames(df_points) <- with(df_points,
                            paste(SHIP_CODE, SAMPLE_AUTO_NUM, 
                                  Time))

sf_points_tf <- st_as_sf(df_points,
                          coords = c("GPS_LON", "GPS_LAT"),
                          crs = "+proj=longlat +ellps=WGS84")

```


### Check Fantasy samples in map     
```{r}
m_fa <- mapview(sf_points_fa, alpha.regions = 0.2)
m_fa

```

### Check Trollfjord samples in map   
```{r}
m_tf <- mapview(sf_points_tf, alpha.regions = 0.2)
m_tf

```


### Check stations in map  
```{r}

df_aqm_stations <- as.data.frame(df_aqm_stations)

rownames(df_aqm_stations) <- df_aqm_stations$StationCode  

sf_points_stations <- st_as_sf(df_aqm_stations,
                       coords = c("Longitude", "Latitude"),
                       crs = "+proj=longlat +ellps=WGS84")

mapview(sf_points_stations)

```

### Stations vs Fantasy sample positions   
Correspondence:  
- VT4 = sample no 21-22
```{r}

m_fa +mapview(sf_points_stations, color = "red") 

```

### Stations vs Trollfjord sample positions   
Correspondence:  
- VT4 = sample no 23-24    
- VT72 = sample no 21-22     
- VT23 = sample no 17-18    
- VT80 = sample no 11-12  
- VT45 = sample no 13-14  
- VT22 = sample no 15-16   
- VR25 = sample no 6-7   
- VR23 = sample no 4-5  
- VR76 = sample no 2-3  

```{r}

m_tf + mapview(sf_points_stations, color = "red") 

```

### Check specific Trollfjord sample number    
- In contrast to Fantasy, there is a bit mess in Trollfjord's SAMPLE_AUTO_NUM  
- They are all over  
```{r}

# Trollfjord
mapview(sf_points_tf %>% filter(SAMPLE_AUTO_NUM == 13), alpha.regions = 0.2)

#
# Check Fantasy in contrast: sample 21 and 22 are always at the right place
#
if (FALSE){
  mapview(sf_points_fa %>% filter(AUTOMATIC_SAMPLE_COUNTER == 21), alpha.regions = 0.2)
  mapview(sf_points_fa %>% filter(AUTOMATIC_SAMPLE_COUNTER == 22), alpha.regions = 0.2)
}

```

### Trollfjord: Get Ferrybox samples in the vicinity of specific lon, lat   
Only for info, not used later   
- For each station, shows SAMPLE_AUTO_NUM with number of data points in paranthesis 
- Shows that we have a problem  
- For some stations, all the closest SAMPLE_AUTO_NUM are two numbers (e.g. 13 and 14 for VT45)
```{r}
#
# Pick trollfjord data from a given box around lon, lat
#
# tol_lon, tol_lat = tolerance in lon,lat direction
#
get_trollfjord_from_box <- function(lon, lat, tol_lon, tol_lat){
  df_ferrybox_sampledata_tf %>%
    filter(
      GPS_LON >= (lon - tol_lon) &
      GPS_LON <= (lon + tol_lon) &
        GPS_LAT >= (lat - tol_lat) &
        GPS_LAT <= (lat + tol_lat))
}

# df_aqm_stations %>%  filter(StationCode == "VT72")
# Test
# get_trollfjord_from_box(5.5877, 62.3066, 0.04, 0.04) %>% xtabs(~SAMPLE_AUTO_NUM, .)

# Test
# df <- with(df_aqm_stations %>%  filter(StationCode == "VT12"),
#            get_trollfjord_from_box(Longitude, Latitude, 0.04, 0.04))

#
# 
summarize_trollfjord_from_box <- function(...){
  df <- get_trollfjord_from_box(...) %>%
    filter(SAMPLE_AUTO_NUM > 0)
  val <- xtabs(~SAMPLE_AUTO_NUM, df)
  stat <- xtabs(~SAMPLE_AUTO_NUM, df) %>% names()
  if (length(val > 0)){
    result <- paste(stat, paste0("(", val, ")")) %>% paste(collapse = ", ")
  } else {
    result <- NA
  }
  result
}
# Test
# summarize_trollfjord_from_box(5.5877, 62.3066, 0.04, 0.04)

#
# Add variable "TF_stations"
#
df_aqm_stations$TF_stations <- df_aqm_stations %>%
  select(Longitude, Latitude) %>%
  as.list() %>%
  pmap_chr(~summarize_trollfjord_from_box(.x, .y, 0.08, 0.04))

df_aqm_stations %>%
  select(StationCode, TF_stations)

```
### Check VT80 + VT23 as example
Only for info, not used later  
```{r}
df <- with(df_aqm_stations %>%  filter(StationCode == "VT80"),
           get_trollfjord_from_box(Longitude, Latitude, 0.04, 0.04))

# df %>% arrange(Time)

df <- with(df_aqm_stations %>%  filter(StationCode == "VR23"),
           get_trollfjord_from_box(Longitude, Latitude, 0.04, 0.04))

# df %>% arrange(Time)

```

## 5. Trollfjord: Get closest station for each ferrybox data point

### Make UTM coordinates   
```{r}
library(sp)
crs_longlat <- "+proj=longlat +ellps=WGS84 +datum=WGS84"
crs_utm <- "+proj=utm +zone=32 +ellps=WGS84 +datum=WGS84 +units=m"

#
# a. Add UTM coor to 'df_ferrybox_sampledata_tf'
#
SP <- SpatialPoints(df_ferrybox_sampledata_tf[,c("GPS_LON", "GPS_LAT")],
         proj4string=CRS(crs_longlat)
         )
SP.UTM <- spTransform(SP, CRS(crs_utm))
# Add transformed coords to data set
df_ferrybox_sampledata_tf$UTM_x <- SP.UTM@coords[,1]
df_ferrybox_sampledata_tf$UTM_y <- SP.UTM@coords[,2]


#
# b. Add UTM coor to 'df_aqm_stations'
#
SP <- SpatialPoints(df_aqm_stations[,c("Longitude", "Latitude")],
         proj4string=CRS(crs_longlat)
         )
SP.UTM <- spTransform(SP, CRS(crs_utm))
# Add transformed coords to data set
df_aqm_stations$UTM_x <- SP.UTM@coords[,1]
df_aqm_stations$UTM_y <- SP.UTM@coords[,2]

plot(UTM_y ~ UTM_x, df_ferrybox_sampledata_tf)
plot(UTM_y ~ UTM_x, df_aqm_stations)
```

### Function  
```{r}
#
# Returns distance to give station (in km)
#
get_distance <- function(station, ferryboxdata){
  df_station <- df_aqm_stations %>% filter(StationCode %in% station)
  result <- ferryboxdata %>%
    mutate(dx = (UTM_x - df_station$UTM_x[1])/1000,
           dy = (UTM_y - df_station$UTM_y[1])/1000,
           Dist = sqrt((dx^2) + (dy^2))
    ) %>%
    select(Dist)
  names(result) <- station
  result
}
# test
# x <- get_distance("VT76", df_ferrybox_sampledata_tf)
# range(x, na.rm = TRUE)
```

### Make distance matrix and find closest station    
```{r}
stations <- df_aqm_stations %>%
  filter(StationCode != "VT4") %>%    # Skagerrak
  pull(StationCode)

distance_matrix <- stations %>% map_dfc(get_distance, ferryboxdata = df_ferrybox_sampledata_tf)

# For test:
# distance_matrix <- distance_matrix[1:4,]
# distance_matrix

distance_minimum <- data.frame(
  Distance_closest = apply(distance_matrix, 1, min),
  Stat_no = apply(distance_matrix, 1, which.min)
)
distance_minimum$StationCode_closest <- stations[distance_minimum$Stat_no]

head(distance_minimum)

if (!"StationCode_closest" %in% names(df_ferrybox_sampledata_tf)){
  df_ferrybox_sampledata_tf <- bind_cols(
    df_ferrybox_sampledata_tf,
    distance_minimum %>% select(StationCode_closest, Distance_closest)
    )
}

#
# This was overkill:
#
# # Extra functions for second lowest number and index of second lowest number
# min.second <- function(x) sort(x)[2]
# which.min.second <- function(x) which(rank(x) == 2)[1]
# # Test
# # min.second(c(3,6,5,9))
# # which.min.second(c(3,6,5,9))
# 
# distance_minimum <- data.frame(
#   Distance1 = apply(distance_matrix, 1, min),
#   Distance2 = apply(distance_matrix, 1, min.second),
#   Stat_no1 = apply(distance_matrix, 1, which.min),
#   Stat_no2 = apply(distance_matrix, 1, which.min.second)
# )
# distance_minimum$StationCode_min1 <- stations[distance_minimum$Stat_no1]
# distance_minimum$StationCode_min2 <- stations[distance_minimum$Stat_no2]

```

### Save 'df_ferrybox_sampledata_tf'    
```{r}

saveRDS(df_ferrybox_sampledata_tf,
        "Datasett/21_df_ferrybox_sampledata_tf_stations.rds")

openxlsx::write.xlsx(df_ferrybox_sampledata_tf,
                     "Datasett/Ferrybox_samples_OneDrive/TF_2018-19_automatic_samples_stations.xlsx")

```



## 6. Check positions of manual samples    
Only for information  

### Check samples in map - prepare data    
```{r}

# Reload the old one
df_ferrybox_sampledata_tf <- readRDS("Datasett/21_df_ferrybox_sampledata_tf.rds")


library(mapview)
library(sf)

#
# Fantasy (sf_points_fa)
#

# Set rownames to "FullStationName" - these will show up when you hover over points with the mouse
df_points <- df_ferrybox_sampledata_fa %>%
  filter(MANUAL_SAMPLE_COUNTER > 0) %>%
  # "copy" coordinates so the show up in popup
  mutate(Lon = GPS_LONGITUDE, Lat = GPS_LATITUDE) %>%  
  as.data.frame()   # read_excel makes a tibble, but you can't set row names for tibbles 

# Set rownames (shown at mouse hover) - must be unique
rownames(df_points) <- with(df_points,
                            paste(SHIP_CODE, MANUAL_SAMPLE_COUNTER, 
                                  SYSTEM_DATE_DMY, SYSTEM_TIME))

sf_points_fa <- st_as_sf(df_points,
                          coords = c("GPS_LONGITUDE", "GPS_LATITUDE"),
                          crs = "+proj=longlat +ellps=WGS84")

#
# Trollfjord (sf_points_tf)
#

# Set rownames to "FullStationName" - these will show up when you hover over points with the mouse
df_points <- df_ferrybox_sampledata_tf %>%
  filter(SAMPLE_MAN_NUM > 0) %>%
  # "copy" coordinates so the show up in popup
  mutate(Lon = GPS_LON, Lat = GPS_LAT) %>%  
  as.data.frame()   # read_excel makes a tibble, but you can't set row names for tibbles 

# Set rownames (shown at mouse hover) - must be unique
rownames(df_points) <- with(df_points,
                            paste(SHIP_CODE, SAMPLE_MAN_NUM, 
                                  SYSTEM_DATE, SYSTEM_TIME))

sf_points_tf <- st_as_sf(df_points,
                          coords = c("GPS_LON", "GPS_LAT"),
                          crs = "+proj=longlat +ellps=WGS84")

```


### Check Fantasy samples in map     
```{r}
m_fa <- mapview(sf_points_fa, alpha.regions = 0.2)
m_fa

```


### Check Trollfjord samples in map     
```{r}
m_fa <- mapview(sf_points_tf, alpha.regions = 0.2)
m_fa

```





