---
title: "22. Match FBdata to AqM stations"
author: "DHJ"
date: "17 February 2020"
output: 
  html_document:
    toc: true    
    toc_flo: true
  github_document:
    toc: true
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_dir = "output", output_format = "all") })    

---

Add Ferrybox data (Temp,Salt) to Aquamonitor sample data
 

## 1. Libraries
```{r, message=FALSE, warning=FALSE}

library(plyr)      # used by functions in "Get_files_NIVA_ftp_server_functions.R" 
library(dplyr)     # load AFTER plyr, so e,g, we use count() from dplyr
library(purrr)
library(ggplot2)
library(lubridate)
library(readxl)
library(tidyr)
library(RColorBrewer)

source("Get_files_NIVA_ftp_server_functions.R")

# library(niRvana)
#source("12_QA_2019_from_excel_functions.R")
# RColorBrewer::display.brewer.all()
```

## 2. Data

### Stations
```{r}
df_aqm_stations <- read_excel("Datasett/AqM_2017_2019_ØKOKYST_Ferrybox_ToR.xlsx", 
                              sheet = "StationPoint")

```

### Ferrybox sample data  
```{r}

df_fa <- readRDS("Datasett/21_df_ferrybox_sampledata_fa.rds")    
df_tf <- readRDS("Datasett/21_df_ferrybox_sampledata_tf_stations.rds")  # has StationCode_closest

```

### Ferrybox: Add Date + Time    
Already did this for Trollfjord (script 21)  
```{r}
df_fa <- df_fa %>% 
  mutate(
    x1 = as.character(SYSTEM_DATE_DMY),
    x2 = as.character(SYSTEM_TIME) %>% substr(12,19),
    Date = ymd(x1),
    Time = ymd_hms(paste(x1, x2))) %>%   # select(x1, x2, Date, Time)
  select(-x1, -x2, -SYSTEM_TIME, -SYSTEM_DATE_DMY) %>%
  select(SHIP_CODE, Date, Time, everything())

df_tf <- df_tf %>%
  select(SHIP_CODE, Date, Time, everything())
  

```

### Ferrybox: add StationCode  
Corresponding numbers in Fantasy (AUTOMATIC_SAMPLE_COUNTER):    
- VT4 = sample no 21-22  
In Trollfjord, these are OFTEN but not always Corresponding numbers (SAMPLE_AUTO_NUM):    
- VT4 = sample no 23-24    
- VT72 = sample no 21-22     
- VT23 = sample no 17-18    
- VT80 = sample no 11-12  
- VT45 = sample no 13-14  
- VT22 = sample no 15-16   
- VR25 = sample no 6-7   
- VR23 = sample no 4-5  
- VT76 = sample no 2-3  
```{r}
df_fa <- df_fa %>%
  mutate(
    StationCode = 
      case_when(
        AUTOMATIC_SAMPLE_COUNTER %in% 21:22 ~ "VT4"
      ))

# Just for interactive use (so one gets a blue little arrow in the envir. window) 
# df_fa <- df_fa %>% filter(TRUE)      
# df_tf <- df_tf %>% filter(TRUE)      

```



### Aquamonitor bottle data
```{r}

# DOC	  Fluo	  KlfA	NH4-N	NO3 	PO4-P	  Salt	SiO2	Temperatur	TOC	TOTN	TOTP	TSM	Turbiditet
# mg/L C	?g/l	?g/l	?g/l	?g/l	?g/l		      mg/l	C	          mg/l ?g/l	?g/l	mg/l	FNU

# sjekke filer i folder "Datasett"
#dir("Datasett")

df_aqm <- read_excel("Datasett/AqM_2017_2019_ØKOKYST_Ferrybox_ToR.xlsx", sheet = "WaterChemistry")

#lage datoobjekt
df_aqm$Date <- as.Date (paste (df_aqm$Year, df_aqm$Month, df_aqm$Day, sep ="-"))

str(df_aqm)
summary(df_aqm)


# no duplicates
str (test <- distinct (df_aqm))

#include only 2018 and 2019 data
df_aqm <- subset(df_aqm, Date >= "2018-01-01" &  Date <= "2019-12-31")

# StationCode seems to be correct for all rows
#levels(as.factor (df_aqm$StationName))
#levels(as.factor (df_aqm$StationCode))
df_aqm$StationCode <- as.factor (df_aqm$StationCode)
df_aqm$StationName <- as.factor (df_aqm$StationName)
levels(df_aqm$StationCode)

#order stations from south to north
df_aqm$StationCode <- factor(df_aqm$StationCode, levels = c("VT4", "VT12", "VT72", "VT80", "VT45", "VT22", "VT23", "VT76", "VR23", "VR25"))

# convert some columns read as chr to numeric
df_aqm<- df_aqm %>% mutate(
    KlfA = as.numeric(KlfA), NH4 = as.numeric(NH4), NO3_NO2 = as.numeric(NO3_NO2),  PO4 = as.numeric(PO4) 
)

# rounding
# df_aqm <- df_aqm %>% 
#   mutate_if(is.numeric, round, digits = 2)

# look at missing values
tb <- df_aqm %>% 
  gather("Variable", Value, DOC:TOTP) %>%
  filter(!is.na(Value)) %>%
  xtabs(~StationCode + Variable, .)
tb

```

### Converting DOC from mg C /l (NIVAlab) to umol C /Kg
```{r}
# https://ocean.ices.dk/Tools/UnitConversion.aspx
#Conversions
#??g/l = ??mol/l ? MW
#g/l ??? g/kg ? 1.025

# converting mg/l to ug/L
DOC_ug <- df_aqm$DOC * 1000

# converting ug/l to umol/L
df_aqm$DOC_umolL <- DOC_ug / 12.01

# converting umol/l to umol/kg
df_aqm$DOC_umolKg <- df_aqm$DOC_umolL / 1.025

```



## 3. Ferrybox data - set StationCode 

### Ferrybox: check StationCode  
Plot the selected Trollfjord samples ('data') in blue +  
the selected fixed stations ('station') in red
```{r}
library(mapview)
library(sf)

#
# Plots the selected Trollfjord samples ('data') in blue +
# the selected fixed stations ('station') in red
#
check_positions_tf <- function(data, station){
  # FB points
  df_points <- data %>%
    # "copy" coordinates so the show up in popup
    mutate(Lon = GPS_LON, Lat = GPS_LAT) %>%  
    as.data.frame()   # read_excel makes a tibble, but you can't set row names for tibbles 
  # Set rownames (shown at mouse hover) - must be unique
  rownames(df_points) <- with(df_points, paste(SHIP_CODE, SAMPLE_AUTO_NUM, Time))
  sf_points_tf <- st_as_sf(df_points,
                           coords = c("GPS_LON", "GPS_LAT"),
                           crs = "+proj=longlat +ellps=WGS84")

  # Stations  
  df_aqm_stations <- df_aqm_stations %>%
    filter(StationCode == station) %>%
    as.data.frame()
  rownames(df_aqm_stations) <- df_aqm_stations$StationCode  
  sf_points_stations <- st_as_sf(df_aqm_stations,
                         coords = c("Longitude", "Latitude"),
                         crs = "+proj=longlat +ellps=WGS84")

  
  mapview(sf_points_tf, alpha.regions = 0.2) + 
    mapview(sf_points_stations, alpha.regions = 0.2, color = "red")

}

# Example
df_tf %>% filter(StationCode_closest %in% "VT76" & Distance_closest < 20) %>% check_positions_tf("VT76")

# Other examples
if (FALSE){
  df_tf %>% filter(StationCode_closest %in% "VT72" & Distance_closest < 10) %>% check_positions_tf("VT72")
  df_tf %>% filter(StationCode_closest %in% "VT23" & Distance_closest < 8) %>% check_positions_tf("VT23")
  df_tf %>% filter(StationCode_closest %in% "VT80" & Distance_closest < 20) %>% check_positions_tf("VT80")
  df_tf %>% filter(StationCode_closest %in% "VT45" & Distance_closest < 20) %>% check_positions_tf("VT45")
  df_tf %>% filter(StationCode_closest %in% "VT22" & Distance_closest < 20) %>% check_positions_tf("VT22")
  # df_tf %>% filter(StationCode_closest %in% "VT25" & Distance_closest < 20) %>% check_positions_tf("VT25")
}

# Other examples II
if (FALSE){
  df_tf %>% 
    filter(StationCode_closest %in% "VT76" & Distance_closest < 4) %>% 
    filter(TRIP == 8488 & year(Date) == 2018) %>% 
    check_positions_tf("VT76")
  
  df_tf %>% 
    filter(StationCode_closest %in% "VT22" & Distance_closest < 4) %>% 
    filter(TRIP == 8486 & year(Date) == 2018) %>% 
    check_positions_tf("VT22")
}



```

### Fix TRIP for Trollfjord  
- The trip number (TRIP) is unreliable, it restarts quite often  
- We make TRIP2 which actiually identifies each trip   
```{r}
# Fantasy - also looks bad
# But we don't care now, seems to work
# ggplot(df_fa, aes(x = Time, y = TRIP_NUMBER)) + geom_point()

ggplot(df_tf, aes(x = Time, y = TRIP)) + geom_point()

df_tf <- df_tf %>%
  arrange(Time) %>%
  ungroup() %>%
  mutate(Restart = ifelse(TRIP < lag(TRIP), 1, 0))
df_tf$Restart[1] <- 0
df_tf$Trip_batch = cumsum(df_tf$Restart)
df_tf <- df_tf %>%
  mutate(TRIP2 = Trip_batch*10000 + TRIP) %>%
  select(-Restart, -Trip_batch)
# df_tf %>% select(TRIP, Restart, Trip_batch)

ggplot(df_tf, aes(x = Time, y = TRIP2)) + geom_point()

```


### Set StationCode for Trollfjord  
```{r}

df_tf <- df_tf %>%
  mutate(
    StationCode = 
      case_when(
        SAMPLE_AUTO_NUM == 0 ~ as.character(NA),
        StationCode_closest %in% "VT72" & Distance_closest < 10 ~ "VT72",
        StationCode_closest %in% "VT23" & Distance_closest <  8 ~ "VT23",
        StationCode_closest %in% "VT80" & Distance_closest < 20 ~ "VT80",
        StationCode_closest %in% "VT45" & Distance_closest < 20 ~ "VT45",
        StationCode_closest %in% "VT22" & Distance_closest < 4 ~ "VT22",
        StationCode_closest %in% "VT76" & Distance_closest < 4 ~ "VT76"
      )) %>%
  mutate(Year = year(Date)) %>%
  group_by(Year, StationCode, TRIP2) %>%
  mutate(Date_range = diff(range(Date)), 
         Time_range = diff(range(Time)),
         n = n()) %>%
  ungroup() %>%
  # Remove StationCode for a single øine of a single station (from plot in next chunk)
  # This had ~4 hours from measurement 2 to 3
  mutate(StationCode = 
           case_when(
             StationCode %in% "VT22" & TRIP2 %in% 8486 & Time_range > 5000 ~ as.character(NA),
             TRUE ~ StationCode))


```

### Time range   
In days and in seconds    
```{r}
#
# Check Date_range - only zero *thumbs up*
# 
df_tf %>% filter(!is.na(StationCode)) %>% xtabs(~Date_range, .)
 
# 
df_tf %>% 
  filter(!is.na(StationCode)) %>% 
  ggplot(aes(Date, Time_range)) + geom_point()

# Checked this one and removed one case - see "special case" in previous chunk
# df_tf %>%
#   filter(StationCode %in% "VT22" & TRIP2 ==	8486)

```


### Check StationCode for Trollfjord
```{r}

if (FALSE){
  df_tf %>%
    xtabs(~paste(year(Date), TRIP) + addNA(StationCode), .)
}

if (FALSE){
  df_tf %>%
    filter(paste(year(Date), TRIP) == "2018 8488" & StationCode == "VT76")
  df_tf %>%
    filter(paste(year(Date), TRIP) == "2018 8486" & StationCode == "VT22")
}
```



## 4. Aggregate ferrybox data to samples and add StationCode   

### Aggregation  
```{r}

span <- function(x, na.rm = TRUE) diff(range(as.numeric(x), na.rm = na.rm))

df_tf_agg <- df_tf %>%
  rename(
    FB_Lon = GPS_LON,
    FB_Lat = GPS_LAT,
    FB_Temp = TEMP_INLET,
    FB_Salt = SAL_CTD,
    FB_Trip = TRIP2) %>%
  group_by(StationCode, FB_Trip) %>%
  summarise_at(vars(Date, Time, FB_Lon, FB_Lat, FB_Temp, FB_Salt), 
               list(mean=mean, range=span), 
               na.rm = TRUE) %>%
  filter(!is.na(StationCode))

df_fa_agg <- df_fa %>%
  mutate(Ferrybox = "fantasy") %>%
  rename(
    FB_Lon = GPS_LONGITUDE,
    FB_Lat = GPS_LATITUDE,
    FB_Temp = INLET_TEMPERATURE,
    FB_Salt = CTD_SALINITY,
    FB_Trip = TRIP_NUMBER) %>%
  group_by(StationCode, FB_Trip) %>%
  summarise_at(vars(Date, Time, FB_Lon, FB_Lat, FB_Temp, FB_Salt), 
               list(mean=mean, range=span), 
               na.rm = TRUE) %>%
  filter(!is.na(StationCode)) %>%
  mutate()

df_ferrybox <- bind_rows(
  df_tf_agg %>% mutate(Ferrybox = "trollfjord"),
  df_fa_agg %>% mutate(Ferrybox = "fantasy")
)

```


### Check Date_range  
If OK (it is), we replace the names Date_mean, Time_man with Date, Time    
```{r}

# Date range
table(df_ferrybox$Date_range)
# ggplot(df_ferrybox, aes(Date_mean, Date_range, color = Ferrybox)) + geom_point()

# Time range
ggplot(df_ferrybox, aes(Date_mean, Time_range, color = Ferrybox)) + geom_point()

#
# Rename these columns
#
df_ferrybox <- df_ferrybox %>%
  rename(Date = Date_mean,
         Time = Time_mean)

```

### Check position range   
Always small  
```{r}

ggplot(df_ferrybox, aes(Date, FB_Lon_range, color = Ferrybox)) + geom_point()
ggplot(df_ferrybox, aes(Date, FB_Lat_range, color = Ferrybox)) + geom_point()

```


## 5. Station positions and dates   
Make "df_station_date"  
```{r}

df_station_pos <- read_excel("Datasett/AqM_2017_2019_ØKOKYST_Ferrybox_ToR.xlsx", sheet = "StationPoint")

df_station_pos$StationCode <- factor(df_station_pos$StationCode, 
                                     levels = c("VT4", "VT12", "VT72", "VT80", "VT45", "VT22", 
                                                "VT23", "VT76", "VR23", "VR25"))

# table(addNA(df_station_pos$StationCode))

my_map <- map_data("world", "Norway")

ggplot(df_station_pos, aes(x = Longitude, y = Latitude)) +
  geom_text(aes(label = StationCode)) +
  annotation_map(my_map, fill = NA, color = "blue") + 
  labs(title = "Stations")

df_station_pos <- df_station_pos %>%
  mutate(Ferrybox = ifelse(StationCode %in% "VT4", "fantasy", "trollfjord"))

# One line per station/date combination in data
df_station_date <- df_aqm %>%
  count(StationCode, Date) %>%
  left_join(subset(df_station_pos, select = c(StationCode, Latitude, Longitude, Ferrybox)),
            by = "StationCode")

# df_station_date
# df_station_date %>% arrange(Date, StationCode)

ggplot(df_station_date, aes(Date, StationCode, color = Ferrybox)) +
  geom_point() +
  geom_vline(xintercept = seq(ymd("2018-01-01"), by = "month", length.out = 24), color = "grey60") +
  labs(title = "Stations/dates of bottle data")

ggplot(df_tf %>% filter(SAMPLE_AUTO_NUM > 0), 
       aes(Date, SAMPLE_AUTO_NUM, color = SHIP_CODE)) +
  geom_point() +
  geom_vline(xintercept = seq(ymd("2018-01-01"), by = "month", length.out = 24), color = "grey60") +
  labs(title = "Stations/dates of all ferrybox 'sample' data")

ggplot(df_tf_agg, aes(Date_mean, StationCode)) +
  geom_point() +
  geom_vline(xintercept = seq(ymd("2018-01-01"), by = "month", length.out = 24), color = "grey60") +
  labs(title = "Stations/dates of bottle data") +
  labs(title = "Stations/dates of Trollfjord ferrybox 'sample' data linked to stations")


```


### Check samples for example months, Trollfjord  
- Bottle data = red points    
- Ferrybox data = open circles  
```{r}

year <- 2018

for (month in 4:8){
  gg <- ggplot(df_station_date %>% 
                 filter(year(Date) == 2018 & month(Date) == month & Ferrybox == "trollfjord"), 
               aes(Date, StationCode)) +
    # Bottle data = red points:
    geom_point(color = "red") +
    # Ferrybox data = open circles:
    geom_point(data = df_ferrybox %>% 
                 filter(year(Date) == 2018 & month(Date) == month & Ferrybox == "trollfjord"),
               shape = 1, size = 3) +
    # Show the entire month:
    scale_x_date(limits = ymd(c(paste0("2018-", month, "-01"),
                                paste0("2018-", month + 1, "-01")
                                ))
    )
  print(gg)
}


```

### Check samples for example months, Fantasy  
- Bottle data = red points    
- Ferrybox data = open circles  
```{r}

year <- 2018
month <- 1:12
gg <- ggplot(df_station_date %>% 
               filter(year(Date) == 2018 & month(Date) %in% month & Ferrybox == "fantasy"), 
             aes(Date, StationCode)) +
  # Bottle data = red points
  geom_point(color = "red") +
  # Ferrybox data = open circles
  geom_point(data = df_ferrybox %>% 
               filter(year(Date) == 2018 & month(Date) %in% month & Ferrybox == "fantasy"),
             shape = 1, size = 3)


```

## 6. Add ferrybox data to Aquamonitor bottle data  
### Function for getting Ferrybox time closest to the given 'time'
```{r}
# Returns a data frame
get_closest_day <- function(date, station, vessel, ferryboxdata){
  ferryboxdata %>%
    filter(StationCode %in% station & Ferrybox %in% vessel) %>%
    mutate(Days_diff = (Date - date)/ddays(1)) %>%                     # difference in number of days
    filter(abs(Days_diff) %in% min(abs(Days_diff), na.rm = TRUE)) %>%  # get the smallest difference
    rename(FB_Date = Date) %>%  # To avoid confusion
    mutate(Bottle_date = date)
    }

# Test
# get_closest_day(ymd("2018-08-28"), "VT22", "trollfjord", df_ferrybox)
# get_closest_day(ymd("2018-01-01"), "VT22", "trollfjord", df_ferrybox)

```


### Get ferrybox data for each of the stations/dates in bottle data   
```{r}

data_for_join <- df_station_date %>%
  count(Date, StationCode, Ferrybox) %>%
  select(-n) %>%
  as.list() %>% 
  purrr::pmap_dfr(~get_closest_day(..1, ..2, ..3, df_ferrybox))

head(data_for_join)

```

### Check difference in days between bottle and ferrybox sample    
Either 0 days difference, or at least 14    
- That means we can just keep the 0 day difference data  
```{r}

ggplot(data_for_join, aes(Days_diff)) + geom_histogram(binwidth = 1)

# xtabs(~Days_diff , data_for_join)

```

### Add to bottle data  
```{r}
# names(data_for_join)
# names(df_aqm)

# Prepare file for 
data_for_join2 <- data_for_join %>%
  filter(abs(Days_diff) <= 1) %>%    # we accpt one dday difference (all are zero in this case)
  rename(FB_Time = Time,
         FB_Lon = FB_Lon_mean,
         FB_Lat = FB_Lat_mean,
         FB_days_diff = Days_diff) %>%
  select(-FB_Trip, -Date_range, -Time_range, -FB_Lon_range, -FB_Lat_range)

if (!"FB_Temp_mean" %in% names(df_aqm)){
  df_aqm <- df_aqm %>%
    mutate(Ferrybox = ifelse(StationCode %in% "VT4", "fantasy", "trollfjord")) %>%
    # Add all ferrybox data
    left_join(data_for_join2, 
              by = c("Ferrybox", "StationCode", "Date" = "Bottle_date"))
}

# Station sequence (along the coast)
# table(addNA(df_aqm$StationCode))

sts <- c("VT4", "VT12", "VT72", "VT23", "VT80", "VT45", "VT22", "VR25", "VR23", "VT76")
# Check
# unique(df_aqm$StationCode) %in% sts

df_aqm <- df_aqm %>%
  mutate(StationCode = factor(StationCode, levels = sts))

```


### Check  
```{r}

df_aqm %>%
  mutate(`Has FB data` = ifelse(is.na(FB_Temp_mean), "No", "Yes")) %>%
  ggplot(aes(Date, StationCode, color = `Has FB data`)) +
  geom_point() +
  scale_color_brewer(palette = "Set1") +
  geom_vline(xintercept = seq(ymd("2018-01-01"), by = "month", length.out = 24), color = "grey60") +
  labs(title = "Stations/dates of bottle data") +
  theme_minimal()

```


## Check ferrybox *log* data  
```{r}

```

